### 2015-02-24 untitled folder 2/app.tex
histograms
histograms

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{0.5cm} | p{7cm} | p{1cm} | c |}
\hline
 Rank & Q &  VUR & Distribution  \\ 
\hline
1 & Learnedthelanguageyouwerespeaking & 2.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedthelanguageyouwerespeakingAPPSERVER} \\ \hline 
2 & Learnedwhenhowmuchandwhatyouareeating & 12.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowmuchandwhatyouareeatingAPPSERVER} \\ \hline 
3 & Learnedyourheartrate & 9.09\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourheartrateAPPSERVER} \\ \hline 
4 & Learnedyourage & 14.29\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourageAPPSERVER} \\ \hline 
5 & Learnedwhenyouarebusyorinterruptible & 17.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouarebusyorinterruptibleAPPSERVER} \\ \hline 
6 & Learnedhowmuchyouuseyourphone & 15.79\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourphoneAPPSERVER} \\ \hline 
7 & Learnedyourfingerprintsomehow & 55.32\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourfingerprintsomehowAPPSERVER} \\ \hline 
8 & Learnedwhereyouarecurrentlygoing & 35.29\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouarecurrentlygoingAPPSERVER} \\ \hline 
9 & Recordedyourworkconversations & 53.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourworkconversationsAPPSERVER} \\ \hline 
10 & Learnedwhoyouwerespendingtimewith & 31.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhoyouwerespendingtimewithAPPSERVER} \\ \hline 
11 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoudownloadedontheinternetsavedonyourdeviceAPPSERVER} \\ \hline 
12 & Learnedwhenyouwerehappyorhavingfun & 24.32\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerehappyorhavingfunAPPSERVER} \\ \hline 
13 & Tookavideoofyouunclothed & 90.63\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouunclothedAPPSERVER} \\ \hline 
14 & Trackedwhereyouare(Likeagps) & 25.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouare(likeaGPS)APPSERVER} \\ \hline 
15 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 38.3\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofpeople(withanoutward-facingcamera)atrandomAPPSERVER} \\ \hline 
16 & Learnedyourcreditcardinformation & 65.52\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourcreditcardinformationAPPSERVER} \\ \hline 
17 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 30.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouareveryaccurately(whereyouareinaroom)APPSERVER} \\ \hline 
18 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 13.79\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowcloseyouaretootherpeopleyouinteractwithAPPSERVER} \\ \hline 
19 & Learnedwhenyouwerelyingnervousorstressed & 41.46\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerelyingnervousorstressedAPPSERVER} \\ \hline 
20 & Copiedanduploadedmusicfromyourdevice & 20.83\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedmusicfromyourdeviceAPPSERVER} \\ \hline 
21 & Sharedvideosotherssentyousavedonyourdevice & 30.56\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosotherssentyousavedonyourdeviceAPPSERVER} \\ \hline 
22 & Learnedyoursocialsecuritynumber & 88.68\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyoursocialsecuritynumberAPPSERVER} \\ \hline 
23 & Learnedwhereyouworksomehow & 46.34\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouworksomehowAPPSERVER} \\ \hline 
24 & Learnedwhatmedicalconditionsyouhave & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicalconditionsyouhaveAPPSERVER} \\ \hline 
25 & Recordedyourpassingconversations & 55.88\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourpassingconversationsAPPSERVER} \\ \hline 
26 & Tookapictureofyoursignature & 59.26\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookapictureofyoursignatureAPPSERVER} \\ \hline 
27 & Learnedhowmuchyouuseyourcomputer & 28.57\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourcomputerAPPSERVER} \\ \hline 
28 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 71.88\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingvideoofyoudoingsomethingembarrassingAPPSERVER} \\ \hline 
29 & Tookavideoofyouoff-Guard & 53.85\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouoff-guardAPPSERVER} \\ \hline 
30 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedaudiorecordingsyoumadeonyourdeviceAPPSERVER} \\ \hline 
31 & Learnedyourusernameforwebsites & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameforwebsitesAPPSERVER} \\ \hline 
32 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 38.1\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyoutalkingtoyourself(makingvoicenotes)APPSERVER} \\ \hline 
33 & Learnedhowmuchmoneyyouhave & 63.41\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchmoneyyouhaveAPPSERVER} \\ \hline 
34 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 18.92\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureAPPSERVER} \\ \hline 
35 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 22.58\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)APPSERVER} \\ \hline 
36 & Recordedyourphoneconversations & 56.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourphoneconversationsAPPSERVER} \\ \hline 
37 & Learnedhowmuchdebtyouhave & 59.46\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchdebtyouhaveAPPSERVER} \\ \hline 
38 & Learnedyourusernameandpasswordforwebsites & 64.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameandpasswordforwebsitesAPPSERVER} \\ \hline 
39 & Learnedyourgender & 9.52\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourgenderAPPSERVER} \\ \hline 
40 & Sharedphotosyoutookwhicharealreadyonyourdevice & 27.27\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoutookwhicharealreadyonyourdeviceAPPSERVER} \\ \hline 
41 & Tookphotosatwork(Withanoutward-Facingcamera) & 62.16\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosatwork(withanoutward-facingcamera)APPSERVER} \\ \hline 
42 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 78.05\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingphotoofyoudoingsomethingembarrassingAPPSERVER} \\ \hline 
43 & Learnedyouraddress & 51.16\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyouraddressAPPSERVER} \\ \hline 
44 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinadigitalpasscodetoalockeddoorAPPSERVER} \\ \hline 
45 & Recordedthesoundaroundyou & 35.9\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedthesoundaroundyouAPPSERVER} \\ \hline 
46 & Tookavideoofyouenteringinyourpinatanatm & 88.57\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinyourPINatanATMAPPSERVER} \\ \hline 
47 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 28.13\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofpeople(withanoutward-facingcamera)atrandomAPPSERVER} \\ \hline 
48 & Learnedwhereyoulivesomehow & 40.63\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyoulivesomehowAPPSERVER} \\ \hline 
49 & Learnedwhattelevisionshowsyouwatch & 11.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhattelevisionshowsyouwatchAPPSERVER} \\ \hline 
50 & Tookaphotoofyouintoxicated & 53.13\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouintoxicatedAPPSERVER} \\ \hline 
51 & Tookvideosatwork(Withanoutward-Facingcamera) & 51.85\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosatwork(withanoutward-facingcamera)APPSERVER} \\ \hline 
52 & Learnedwhenandhowmuchyouspendtimealone & 35.48\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouspendtimealoneAPPSERVER} \\ \hline 
53 & Learnedwhenyouaresickorhealthy & 26.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouaresickorhealthyAPPSERVER} \\ \hline 
54 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 6.98\% & \includegraphics[width = 2cm, height = 0.5cm]{../scannedyoureyetolearnyoureyepatterns(foreyetracking)APPSERVER} \\ \hline 
55 & Learnedwhatmedicationyoutake & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicationyoutakeAPPSERVER} \\ \hline 
56 & Learnedwhatwebsitesyougoto & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatwebsitesyougotoAPPSERVER} \\ \hline 
57 & Learnedwhenandhowmuchyouhavesex & 73.08\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouhavesexAPPSERVER} \\ \hline 
58 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 66.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)atrandomAPPSERVER} \\ \hline 
59 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)athomeAPPSERVER} \\ \hline 
60 & Tookaphotoofyouoff-Guard & 45.95\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouoff-guardAPPSERVER} \\ \hline 
61 & Learnedyourname & 32.56\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyournameAPPSERVER} \\ \hline 
62 & Learnedwhenandhowwellyouaresleeping & 21.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowwellyouaresleepingAPPSERVER} \\ \hline 
63 & Tookaphotoofyouunclothed & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouunclothedAPPSERVER} \\ \hline 
64 & Tookavideoofyouintoxicated & 68.75\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouintoxicatedAPPSERVER} \\ \hline 
65 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 60.53\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)athomeAPPSERVER} \\ \hline 
66 & Learnedyourbankaccountinformation & 90.91\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourbankaccountinformationAPPSERVER} \\ \hline 
67 & Sharedphotosotherssenttoyousavedonyourdevice & 41.18\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosotherssenttoyousavedonyourdeviceAPPSERVER} \\ \hline 
68 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 32.14\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceAPPSERVER} \\ \hline 
69 & Learnedwhenhowandhowmuchyouexercise & 16.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowandhowmuchyouexerciseAPPSERVER} \\ \hline 
70 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 48.39\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)atrandomAPPSERVER} \\ \hline 
71 & Learnedwhatproductsyoubuy & 26.32\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatproductsyoubuyAPPSERVER} \\ \hline 
72 & Sharedvideosyoutookwhicharealreadyonyourdevice & 51.61\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoutookwhicharealreadyonyourdeviceAPPSERVER} \\ \hline 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}