### 2015-02-24 tables/benefit.tex
linda's push

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{2cm} | p{1cm} | p{1cm} | p{1cm} | c |}
\hline
Technology & Q1 &  Median & Q3 & Distribution  \\ 
\hline
Language Detection  & 15.0 & 35.0 & 60.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/languagedetectionben} \\ 
Email  & 29.0 & 50.0 & 77.5 & \includegraphics[width = 2cm, height = 0.5cm]{tables/emailben} \\ 
Heart Rate Detection  & 26.0 & 40.0 & 65.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/heartratedetectionben} \\ 
Cubetastic  & 10.0 & 15.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/cubetasticben} \\ 
Speech To Text  & 15.0 & 25.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/speechtotextben} \\ 
Location Tracking  & 20.0 & 40.0 & 70.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/locationtrackingben} \\ 
Facial Detection  & 10.0 & 20.0 & 34.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/facialdetectionben} \\ 
Age Detection  & 10.0 & 12.0 & 22.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/agedetectionben} \\ 
Discreet Video Camera  & 15.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/discreetvideocameraben} \\ 
Facial Recognition  & 12.5 & 22.0 & 42.5 & \includegraphics[width = 2cm, height = 0.5cm]{tables/facialrecognitionben} \\ 
Internet  & 45.0 & 65.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/internetben} \\ 
Voice Recognition  & 15.0 & 25.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/voicerecognitionben} \\ 
Discreet Microphone  & 10.0 & 15.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/discreetmicrophoneben} \\ 
Motorcycle & 12.0 & 20.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/MotorcycleBenefit} \\ 
Gender Detection  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/genderdetectionben} \\ 
Voice Based Emotion Detection  & 10.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/voicebasedemotiondetectionben} \\ 
Smartwatches  & 10.0 & 20.0 & 35.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/smartwatchesben} \\ 
Electricity & 50.0 & 88.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/ElectricityBenefit} \\ 
Google Glass  & 12.0 & 20.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/googleglassben} \\ 
Laptops  & 40.0 & 60.0 & 80.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/laptopsben} \\ 
Lawnmower & 15.0 & 24.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/LawnmowerBenefit} \\ 
Handgun & 10.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/HandgunBenefit} \\ 
Smartphones  & 30.0 & 50.0 & 75.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/smartphonesben} \\ 
Fitness Trackers  & 10.0 & 18.5 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/fitnesstrackersben} \\ 
\hline
\end{tabular}
\caption{Benefit Table}
\label{top10}
\end{center}
\end{table}