### 2015-02-24 tables/collapsed.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{0.5cm} | p{7cm} | p{1cm} | c |}
\hline
 Rank & Q &  VUR & Distribution  \\ 
\hline
1 & Learnedyourheartrate & 27.5 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourheartratecombined} \\ 
2 & Learnedwhenyouarebusyorinterruptible & 29.5 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenyouarebusyorinterruptiblecombined} \\ 
3 & Tookavideoofyouunclothed & 95.97 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookavideoofyouunclothedcombined} \\ 
4 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 81.05 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)athomecombined} \\ 
5 & Learnedwhenhowandhowmuchyouexercise & 38.66 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenhowandhowmuchyouexercisecombined} \\ 
6 & Learnedwhenyouwerehappyorhavingfun & 34.75 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenyouwerehappyorhavingfuncombined} \\ 
7 & Learnedyoursocialsecuritynumber & 94.84 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyoursocialsecuritynumbercombined} \\ 
8 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 91.39 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookanincriminatingphotoofyoudoingsomethingembarrassingcombined} \\ 
9 & Learnedwhatwebsitesyougoto & 66.44 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhatwebsitesyougotocombined} \\ 
10 & Recordedyourphoneconversations & 85.16 & \includegraphics[width = 0.5cm, height = 0.5cm]{../recordedyourphoneconversationscombined} \\ 
11 & Learnedyourgender & 15.0 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourgendercombined} \\ 
12 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 88.41 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookanincriminatingvideoofyoudoingsomethingembarrassingcombined} \\ 
13 & Learnedyourfingerprintsomehowwith & 77.78 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourfingerprintsomehowwithcombined} \\ 
14 & Recordedyourpassingconversations & 78.13 & \includegraphics[width = 0.5cm, height = 0.5cm]{../recordedyourpassingconversationscombined} \\ 
15 & Learnedhowmuchmoneyyouhave & 84.44 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedhowmuchmoneyyouhavecombined} \\ 
16 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 63.51 & \includegraphics[width = 0.5cm, height = 0.5cm]{../trackedwhereyouareveryaccurately(whereyouareinaroom)combined} \\ 
17 & Sharedphotosyoutookwhicharealreadyonyourdevice & 66.21 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedphotosyoutookwhicharealreadyonyourdevicecombined} \\ 
18 & Tookavideoofyouenteringinyourpinatanatm & 92.67 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookavideoofyouenteringinyourPINatanATMcombined} \\ 
19 & Recordedyourworkconversations & 86.82 & \includegraphics[width = 0.5cm, height = 0.5cm]{../recordedyourworkconversationscombined} \\ 
20 & Tookaphotoofyouintoxicated & 78.95 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookaphotoofyouintoxicatedcombined} \\ 
21 & Learnedwhereyouarecurrentlygoing & 60.87 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhereyouarecurrentlygoingcombined} \\ 
22 & Learnedwhenyouwerelyingnervousorstressed & 69.74 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenyouwerelyingnervousorstressedcombined} \\ 
23 & Learnedwhereyouworksomehow & 58.09 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhereyouworksomehowcombined} \\ 
24 & Learnedyourusernameforwebsites & 73.44 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourusernameforwebsitescombined} \\ 
25 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 61.94 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookphotosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
26 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 71.81 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedphotosyoudownloadedfromtheinternetsavedonyourdevicecombined} \\ 
27 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 61.49 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedvideosyoudownloadedontheinternetsavedonyourdevicecombined} \\ 
28 & Tookavideoofyouintoxicated & 83.21 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookavideoofyouintoxicatedcombined} \\ 
29 & Learnedthelanguageyouwerespeaking & 15.86 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedthelanguageyouwerespeakingcombined} \\ 
30 & Tookavideoofyouoff-Guard & 76.0 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookavideoofyouoff-guardcombined} \\ 
31 & Learnedwhatproductsyoubuy & 50.0 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhatproductsyoubuycombined} \\ 
32 & Learnedwhatmedicalconditionsyouhave & 77.7 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhatmedicalconditionsyouhavecombined} \\ 
33 & Sharedphotosotherssenttoyousavedonyourdevice & 71.63 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedphotosotherssenttoyousavedonyourdevicecombined} \\ 
34 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 40.51 & \includegraphics[width = 0.5cm, height = 0.5cm]{../scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\ 
35 & Learnedwhoyouwerespendingtimewith & 60.0 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhoyouwerespendingtimewithcombined} \\ 
36 & Learnedwhenyouaresickorhealthy & 48.17 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenyouaresickorhealthycombined} \\ 
37 & Learnedyouraddress & 72.61 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyouraddresscombined} \\ 
38 & Learnedwhereyoulivesomehow & 65.0 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhereyoulivesomehowcombined} \\ 
39 & Learnedyourage & 24.29 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyouragecombined} \\ 
40 & Tookphotosatwork(Withanoutward-Facingcamera) & 74.62 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookphotosatwork(withanoutward-facingcamera)combined} \\ 
41 & Learnedyourfingerprintsomehow & 67.23 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourfingerprintsomehowcombined} \\ 
42 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 54.4 & \includegraphics[width = 0.5cm, height = 0.5cm]{../copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)combined} \\ 
43 & Learnedyourusernameandpasswordforwebsites & 89.55 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourusernameandpasswordforwebsitescombined} \\ 
44 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.55 & \includegraphics[width = 0.5cm, height = 0.5cm]{../copiedanduploadedaudiorecordingsyoumadeonyourdevicecombined} \\ 
45 & Tookaphotoofyouoff-Guard & 72.55 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookaphotoofyouoff-guardcombined} \\ 
46 & Tookaphotoofyouunclothed & 92.59 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookaphotoofyouunclothedcombined} \\ 
47 & Learnedwhenandhowwellyouaresleeping & 40.56 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenandhowwellyouaresleepingcombined} \\ 
48 & Learnedyourcreditcardinformation & 88.98 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourcreditcardinformationcombined} \\ 
49 & Copiedanduploadedmusicfromyourdevice & 28.06 & \includegraphics[width = 0.5cm, height = 0.5cm]{../copiedanduploadedmusicfromyourdevicecombined} \\ 
50 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 76.19 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)atrandomcombined} \\ 
51 & Learnedyourname & 42.54 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyournamecombined} \\ 
52 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 46.98 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedhowcloseyouaretootherpeopleyouinteractwithcombined} \\ 
53 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 78.76 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)atrandomcombined} \\ 
54 & Sharedvideosotherssentyousavedonyourdevice & 70.59 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedvideosotherssentyousavedonyourdevicecombined} \\ 
55 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 61.04 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookvideosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
56 & Trackedwhereyouare(Likeagps) & 61.24 & \includegraphics[width = 0.5cm, height = 0.5cm]{../trackedwhereyouare(likeaGPS)combined} \\ 
57 & Learnedhowmuchyouuseyourphone & 51.95 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedhowmuchyouuseyourphonecombined} \\ 
58 & Recordedthesoundaroundyou & 60.45 & \includegraphics[width = 0.5cm, height = 0.5cm]{../recordedthesoundaroundyoucombined} \\ 
59 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 46.81 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturecombined} \\ 
60 & Learnedwhenandhowmuchyouhavesex & 81.95 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenandhowmuchyouhavesexcombined} \\ 
61 & Learnedwhenhowmuchandwhatyouareeating & 42.86 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenhowmuchandwhatyouareeatingcombined} \\ 
62 & Learnedwhatmedicationyoutake & 69.49 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhatmedicationyoutakecombined} \\ 
63 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 69.59 & \includegraphics[width = 0.5cm, height = 0.5cm]{../recordedyoutalkingtoyourself(makingvoicenotes)combined} \\ 
64 & Tookvideosatwork(Withanoutward-Facingcamera) & 70.54 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookvideosatwork(withanoutward-facingcamera)combined} \\ 
65 & Learnedwhattelevisionshowsyouwatch & 30.2 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhattelevisionshowsyouwatchcombined} \\ 
66 & Learnedyourbankaccountinformation & 95.91 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedyourbankaccountinformationcombined} \\ 
67 & Learnedhowmuchyouuseyourcomputer & 44.93 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedhowmuchyouuseyourcomputercombined} \\ 
68 & Learnedwhenandhowmuchyouspendtimealone & 61.27 & \includegraphics[width = 0.5cm, height = 0.5cm]{../learnedwhenandhowmuchyouspendtimealonecombined} \\ 
69 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 85.53 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookavideoofyouenteringinadigitalpasscodetoalockeddoorcombined} \\ 
70 & Sharedvideosyoutookwhicharealreadyonyourdevice & 68.89 & \includegraphics[width = 0.5cm, height = 0.5cm]{../sharedvideosyoutookwhicharealreadyonyourdevicecombined} \\ 
71 & Tookapictureofyoursignature & 68.07 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookapictureofyoursignaturecombined} \\ 
72 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 87.5 & \includegraphics[width = 0.5cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)athomecombined} \\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}