### 2015-02-24 tables/risk.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{2cm} | p{1cm} | p{1cm} | p{1cm} | c |}
\hline
Technology & Q1 &  Median & Q3 & Distribution  \\ 
\hline
Voice Based Emotion Detection  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/voicebasedemotiondetectionrisk} \\ 
Facial Detection  & 10.0 & 10.0 & 25.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/facialdetectionrisk} \\ 
Facial Recognition  & 10.0 & 17.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/facialrecognitionrisk} \\ 
Language Detection  & 10.0 & 10.0 & 10.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/languagedetectionrisk} \\ 
Gender Detection  & 10.0 & 10.0 & 13.5 & \includegraphics[width = 2cm, height = 0.5cm]{tables/genderdetectionrisk} \\ 
Heart Rate Detection  & 10.0 & 10.0 & 10.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/heartratedetectionrisk} \\ 
Age Detection  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/agedetectionrisk} \\ 
Smartwatches  & 10.0 & 10.0 & 10.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/smartwatchesrisk} \\ 
Discreet Microphone  & 10.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/discreetmicrophonerisk} \\ 
Location Tracking  & 10.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/locationtrackingrisk} \\ 
Electricity & 15.0 & 25.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/ElectricityRisk} \\ 
Speech To Text  & 10.0 & 10.0 & 10.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/speechtotextrisk} \\ 
Google Glass  & 10.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/googleglassrisk} \\ 
Laptops  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/laptopsrisk} \\ 
Email  & 10.0 & 10.0 & 18.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/emailrisk} \\ 
Internet  & 10.0 & 15.0 & 31.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/internetrisk} \\ 
Smartphones  & 10.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/smartphonesrisk} \\ 
Discreet Video Camera  & 10.0 & 12.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/discreetvideocamerarisk} \\ 
Cubetastic  & 10.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/cubetasticrisk} \\ 
Voice Recognition  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/voicerecognitionrisk} \\ 
Handgun & 40.0 & 60.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/HandgunRisk} \\ 
Lawnmower & 12.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/LawnmowerRisk} \\ 
Motorcycle & 27.0 & 45.0 & 70.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/MotorcycleRisk} \\ 
Fitness Trackers  & 10.0 & 10.0 & 10.0 & \includegraphics[width = 2cm, height = 0.5cm]{tables/fitnesstrackersrisk} \\ 
\hline
\end{tabular}
\caption{Risk Table}
\label{top10}
\end{center}
\end{table}