### 2015-02-24 tables/work.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{3}{|r|}{{Continued on next page}} \\ \hline \hline
\endfoot
\hline 
\endlastfoot

\hline
1 & Learnedwhenyouwerehappyorhavingfun & 43.75\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfunWORKCONTACTS} \\ \hline 
2 & Recordedyourpassingconversations & 87.5\% & \includegraphics[width = 0.5cm, height = 0.5cm]{recordedyourpassingconversationsWORKCONTACTS} \\ \hline 
3 & Tookavideoofyouenteringinyourpinatanatm & 87.5\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMWORKCONTACTS} \\ \hline 
4 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 61.11\% & \includegraphics[width = 0.5cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)WORKCONTACTS} \\ \hline 
5 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 93.55\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomeWORKCONTACTS} \\ \hline 
6 & Learnedyourgender & 15.79\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourgenderWORKCONTACTS} \\ \hline 
7 & Learnedhowmuchyouuseyourphone & 68.75\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowmuchyouuseyourphoneWORKCONTACTS} \\ \hline 
8 & Learnedwhenyouwerelyingnervousorstressed & 91.89\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenyouwerelyingnervousorstressedWORKCONTACTS} \\ \hline 
9 & Learnedwhenyouaresickorhealthy & 62.22\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenyouaresickorhealthyWORKCONTACTS} \\ \hline 
10 & Tookapictureofyoursignature & 85.71\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookapictureofyoursignatureWORKCONTACTS} \\ \hline 
11 & Learnedyourname & 26.47\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyournameWORKCONTACTS} \\ \hline 
12 & Trackedwhereyouare(Likeagps) & 63.64\% & \includegraphics[width = 0.5cm, height = 0.5cm]{trackedwhereyouare(likeaGPS)WORKCONTACTS} \\ \hline 
13 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 88.24\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)atrandomWORKCONTACTS} \\ \hline 
14 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 89.19\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)athomeWORKCONTACTS} \\ \hline 
15 & Learnedwhereyouarecurrentlygoing & 65.85\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhereyouarecurrentlygoingWORKCONTACTS} \\ \hline 
16 & Tookaphotoofyouintoxicated & 87.1\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookaphotoofyouintoxicatedWORKCONTACTS} \\ \hline 
17 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 78.13\% & \includegraphics[width = 0.5cm, height = 0.5cm]{trackedwhereyouareveryaccurately(whereyouareinaroom)WORKCONTACTS} \\ \hline 
18 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 86.27\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceWORKCONTACTS} \\ \hline 
19 & Learnedwhenhowandhowmuchyouexercise & 66.67\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexerciseWORKCONTACTS} \\ \hline 
20 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 66.67\% & \includegraphics[width = 0.5cm, height = 0.5cm]{copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)WORKCONTACTS} \\ \hline 
21 & Recordedyourphoneconversations & 90.32\% & \includegraphics[width = 0.5cm, height = 0.5cm]{recordedyourphoneconversationsWORKCONTACTS} \\ \hline 
22 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 94.29\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingWORKCONTACTS} \\ \hline 
23 & Learnedyourusernameandpasswordforwebsites & 94.29\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitesWORKCONTACTS} \\ \hline 
24 & Learnedyourheartrate & 44.12\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourheartrateWORKCONTACTS} \\ \hline 
25 & Learnedwhatmedicationyoutake & 81.25\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhatmedicationyoutakeWORKCONTACTS} \\ \hline 
26 & Learnedwhereyoulivesomehow & 69.44\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhereyoulivesomehowWORKCONTACTS} \\ \hline 
27 & Copiedanduploadedmusicfromyourdevice & 42.11\% & \includegraphics[width = 0.5cm, height = 0.5cm]{copiedanduploadedmusicfromyourdeviceWORKCONTACTS} \\ \hline 
28 & Sharedvideosyoutookwhicharealreadyonyourdevice & 79.31\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedvideosyoutookwhicharealreadyonyourdeviceWORKCONTACTS} \\ \hline 
29 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 71.88\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookvideosofpeople(withanoutward-facingcamera)atrandomWORKCONTACTS} \\ \hline 
30 & Learnedhowmuchmoneyyouhave & 93.1\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowmuchmoneyyouhaveWORKCONTACTS} \\ \hline 
31 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 51.35\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowcloseyouaretootherpeopleyouinteractwithWORKCONTACTS} \\ \hline 
32 & Tookavideoofyouoff-Guard & 91.43\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookavideoofyouoff-guardWORKCONTACTS} \\ \hline 
33 & Learnedyourfingerprintsomehow & 70.27\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourfingerprintsomehowWORKCONTACTS} \\ \hline 
34 & Learnedyouraddress & 81.58\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyouraddressWORKCONTACTS} \\ \hline 
35 & Learnedwhenandhowwellyouaresleeping & 62.07\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenandhowwellyouaresleepingWORKCONTACTS} \\ \hline 
36 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 82.86\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedvideosyoudownloadedontheinternetsavedonyourdeviceWORKCONTACTS} \\ \hline 
37 & Tookvideosatwork(Withanoutward-Facingcamera) & 70.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookvideosatwork(withanoutward-facingcamera)WORKCONTACTS} \\ \hline 
38 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 81.48\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)atrandomWORKCONTACTS} \\ \hline 
39 & Tookphotosatwork(Withanoutward-Facingcamera) & 81.48\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookphotosatwork(withanoutward-facingcamera)WORKCONTACTS} \\ \hline 
40 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 63.16\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureWORKCONTACTS} \\ \hline 
41 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 82.05\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookphotosofpeople(withanoutward-facingcamera)atrandomWORKCONTACTS} \\ \hline 
42 & Learnedwhenhowmuchandwhatyouareeating & 75.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenhowmuchandwhatyouareeatingWORKCONTACTS} \\ \hline 
43 & Sharedphotosyoutookwhicharealreadyonyourdevice & 79.55\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedphotosyoutookwhicharealreadyonyourdeviceWORKCONTACTS} \\ \hline 
44 & Learnedwhatwebsitesyougoto & 86.84\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhatwebsitesyougotoWORKCONTACTS} \\ \hline 
45 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 88.57\% & \includegraphics[width = 0.5cm, height = 0.5cm]{recordedyoutalkingtoyourself(makingvoicenotes)WORKCONTACTS} \\ \hline 
46 & Learnedthelanguageyouwerespeaking & 28.13\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedthelanguageyouwerespeakingWORKCONTACTS} \\ \hline 
47 & Tookaphotoofyouoff-Guard & 80.49\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookaphotoofyouoff-guardWORKCONTACTS} \\ \hline 
48 & Sharedvideosotherssentyousavedonyourdevice & 80.49\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedvideosotherssentyousavedonyourdeviceWORKCONTACTS} \\ \hline 
49 & Learnedwhoyouwerespendingtimewith & 76.19\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhoyouwerespendingtimewithWORKCONTACTS} \\ \hline 
50 & Learnedwhenandhowmuchyouhavesex & 90.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenandhowmuchyouhavesexWORKCONTACTS} \\ \hline 
51 & Recordedthesoundaroundyou & 75.76\% & \includegraphics[width = 0.5cm, height = 0.5cm]{recordedthesoundaroundyouWORKCONTACTS} \\ \hline 
52 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 94.59\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingWORKCONTACTS} \\ \hline 
53 & Learnedwhatproductsyoubuy & 62.5\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhatproductsyoubuyWORKCONTACTS} \\ \hline 
54 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 81.08\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookavideoofyouenteringinadigitalpasscodetoalockeddoorWORKCONTACTS} \\ \hline 
55 & Recordedyourworkconversations & 100.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{recordedyourworkconversationsWORKCONTACTS} \\ \hline 
56 & Tookaphotoofyouunclothed & 100.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookaphotoofyouunclothedWORKCONTACTS} \\ \hline 
57 & Tookavideoofyouunclothed & 100.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookavideoofyouunclothedWORKCONTACTS} \\ \hline 
58 & Learnedyourbankaccountinformation & 100.0\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourbankaccountinformationWORKCONTACTS} \\ \hline 
59 & Learnedhowmuchdebtyouhave & 87.18\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowmuchdebtyouhaveWORKCONTACTS} \\ \hline 
60 & Learnedhowmuchyouuseyourcomputer & 45.45\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedhowmuchyouuseyourcomputerWORKCONTACTS} \\ \hline 
61 & Learnedyourusernameforwebsites & 84.62\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourusernameforwebsitesWORKCONTACTS} \\ \hline 
62 & Learnedwhenandhowmuchyouspendtimealone & 80.95\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenandhowmuchyouspendtimealoneWORKCONTACTS} \\ \hline 
63 & Learnedyourcreditcardinformation & 95.12\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourcreditcardinformationWORKCONTACTS} \\ \hline 
64 & Learnedwhenyouarebusyorinterruptible & 36.36\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptibleWORKCONTACTS} \\ \hline 
65 & Learnedwhattelevisionshowsyouwatch & 36.36\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchWORKCONTACTS} \\ \hline 
66 & Learnedyoursocialsecuritynumber & 93.75\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyoursocialsecuritynumberWORKCONTACTS} \\ \hline 
67 & Learnedyourage & 36.67\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedyourageWORKCONTACTS} \\ \hline 
68 & Learnedwhatmedicalconditionsyouhave & 85.29\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhatmedicalconditionsyouhaveWORKCONTACTS} \\ \hline 
69 & Sharedphotosotherssenttoyousavedonyourdevice & 75.61\% & \includegraphics[width = 0.5cm, height = 0.5cm]{sharedphotosotherssenttoyousavedonyourdeviceWORKCONTACTS} \\ \hline 
70 & Learnedwhereyouworksomehow & 57.14\% & \includegraphics[width = 0.5cm, height = 0.5cm]{learnedwhereyouworksomehowWORKCONTACTS} \\ \hline 
71 & Tookavideoofyouintoxicated & 88.89\% & \includegraphics[width = 0.5cm, height = 0.5cm]{tookavideoofyouintoxicatedWORKCONTACTS} \\ \hline 
72 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.22\% & \includegraphics[width = 0.5cm, height = 0.5cm]{copiedanduploadedaudiorecordingsyoumadeonyourdeviceWORKCONTACTS} \\ \hline 
\hline

\end{longtable}
\end{document}