### 2015-02-25 tex-inputs/background.tex
started background and scoping; finished up a couple sections

\section{Current Wearable Devices}  %aka background
%LL: This needs to be a lot better, and sound not so hand-wavey.
To obtain a comprehensive list of possible risks that wearable devices might present in the future, we examined the sensors and capabilities of the most popular wearable devices on the market at the time of this study. The most common wearables of the time were in the form of fitness trackers and smart watches. However, there are also wearable glasses \footnote{ https://www.google.com/intl/en/glass/start/how-it-looks/, http://www.sony.net/SonyInfo/News/Press/201412/14-118E/index.html}, rings \footnote{https://ringly.com/}, bracelets \footnote{http://www.intel.com/content/www/us/en/wearables/fashion-technology.html} and more.

 At the time of this study, August 2014, the most popular wearable devices included the fitbit fitness tracker which performs continuously monitors heartbeat, steps taken, and sleep patterns \footnote{http://www.fitbit.com/, http://time.com/516/26-fitness-trackers-ranked-from-worst-to-first/}, the pebble smartwatch which can take pictures, send texts, show notifications from online, and push notifications to services \footnote{https://getpebble.com/, http://www.theverge.com/2014/1/7/5284144/the-best-wearables-of-ces-2014-stream, http://readwrite.com/2014/02/10/10-cool-things-a-pebble-smartwatch-can-do}, and google glass \footnote{http://en.wikipedia.org/wiki/Google\_Glass, http://www.engadget.com/2014/07/30/the-top-9-wearables-on-you-can-buy-right-now/}. These wearable devices, along with other comparable wearable devices on the market, were researched as inspiration for the survey questions.

 Google glass has since disappeared due to complaints from the public \footnote{http://www.google.com/glass/start/, http://uk.pcmag.com/wearable-tech/36416/opinion/rest-in-peace-google-glass-2012-2014}, and many instances of assault for wearing the product had contributed to this \footnote{http://www.businessinsider.com/i-was-assaulted-for-wearing-google-glass-2014-4, http://www.cnet.com/news/journalist-i-was-assaulted-on-street-for-wearing-google-glass/, http://mashable.com/2014/02/26/google-glass-assault/, http://www.cnn.com/2014/04/14/tech/mobile/google-glass-attack/}.