### 2015-02-26 Tables(2)/app.tex
fixed tables in tables(2)
title

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Learnedyourbankaccountinformation & 90.91\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourbankaccountinformationAPPSERVER} \\ 
2 & Tookavideoofyouunclothed & 90.63\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouunclothedAPPSERVER} \\ 
3 & Learnedyoursocialsecuritynumber & 88.68\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyoursocialsecuritynumberAPPSERVER} \\ 
4 & Tookavideoofyouenteringinyourpinatanatm & 88.57\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMAPPSERVER} \\ 
5 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 78.05\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingAPPSERVER} \\ 
6 & Tookaphotoofyouunclothed & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouunclothedAPPSERVER} \\ 
7 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinadigitalpasscodetoalockeddoorAPPSERVER} \\ 
8 & Learnedwhenandhowmuchyouhavesex & 73.08\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouhavesexAPPSERVER} \\ 
9 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 71.88\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingAPPSERVER} \\ 
10 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomeAPPSERVER} \\ 
11 & Tookavideoofyouintoxicated & 68.75\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouintoxicatedAPPSERVER} \\ 
12 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 66.67\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)atrandomAPPSERVER} \\ 
13 & Learnedyourcreditcardinformation & 65.52\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourcreditcardinformationAPPSERVER} \\ 
14 & Learnedyourusernameandpasswordforwebsites & 64.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitesAPPSERVER} \\ 
15 & Learnedhowmuchmoneyyouhave & 63.41\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchmoneyyouhaveAPPSERVER} \\ 
16 & Tookphotosatwork(Withanoutward-Facingcamera) & 62.16\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosatwork(withanoutward-facingcamera)APPSERVER} \\ 
17 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 60.53\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)athomeAPPSERVER} \\ 
18 & Learnedhowmuchdebtyouhave & 59.46\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchdebtyouhaveAPPSERVER} \\ 
19 & Tookapictureofyoursignature & 59.26\% & \includegraphics[width = 2cm, height = 0.5cm]{tookapictureofyoursignatureAPPSERVER} \\ 
20 & Recordedyourphoneconversations & 56.67\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourphoneconversationsAPPSERVER} \\ 
21 & Recordedyourpassingconversations & 55.88\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourpassingconversationsAPPSERVER} \\ 
22 & Learnedyourfingerprintsomehow & 55.32\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourfingerprintsomehowAPPSERVER} \\ 
23 & Tookavideoofyouoff-Guard & 53.85\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouoff-guardAPPSERVER} \\ 
24 & Recordedyourworkconversations & 53.33\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourworkconversationsAPPSERVER} \\ 
25 & Tookaphotoofyouintoxicated & 53.13\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouintoxicatedAPPSERVER} \\ 
26 & Tookvideosatwork(Withanoutward-Facingcamera) & 51.85\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosatwork(withanoutward-facingcamera)APPSERVER} \\ 
27 & Sharedvideosyoutookwhicharealreadyonyourdevice & 51.61\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoutookwhicharealreadyonyourdeviceAPPSERVER} \\ 
28 & Learnedyouraddress & 51.16\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouraddressAPPSERVER} \\ 
29 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedaudiorecordingsyoumadeonyourdeviceAPPSERVER} \\ 
30 & Learnedyourusernameforwebsites & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameforwebsitesAPPSERVER} \\ 
31 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 48.39\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)atrandomAPPSERVER} \\ 
32 & Learnedwhereyouworksomehow & 46.34\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouworksomehowAPPSERVER} \\ 
33 & Tookaphotoofyouoff-Guard & 45.95\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouoff-guardAPPSERVER} \\ 
34 & Learnedwhenyouwerelyingnervousorstressed & 41.46\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerelyingnervousorstressedAPPSERVER} \\ 
35 & Sharedphotosotherssenttoyousavedonyourdevice & 41.18\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosotherssenttoyousavedonyourdeviceAPPSERVER} \\ 
36 & Learnedwhereyoulivesomehow & 40.63\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyoulivesomehowAPPSERVER} \\ 
37 & Learnedwhatmedicalconditionsyouhave & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicalconditionsyouhaveAPPSERVER} \\ 
38 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 38.3\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofpeople(withanoutward-facingcamera)atrandomAPPSERVER} \\ 
39 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 38.1\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyoutalkingtoyourself(makingvoicenotes)APPSERVER} \\ 
40 & Learnedwhatmedicationyoutake & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicationyoutakeAPPSERVER} \\ 
41 & Learnedwhatwebsitesyougoto & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatwebsitesyougotoAPPSERVER} \\ 
42 & Recordedthesoundaroundyou & 35.9\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedthesoundaroundyouAPPSERVER} \\ 
43 & Learnedwhenandhowmuchyouspendtimealone & 35.48\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouspendtimealoneAPPSERVER} \\ 
44 & Learnedwhereyouarecurrentlygoing & 35.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouarecurrentlygoingAPPSERVER} \\ 
45 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoudownloadedontheinternetsavedonyourdeviceAPPSERVER} \\ 
46 & Learnedyourname & 32.56\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyournameAPPSERVER} \\ 
47 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 32.14\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceAPPSERVER} \\ 
48 & Learnedwhoyouwerespendingtimewith & 31.43\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhoyouwerespendingtimewithAPPSERVER} \\ 
49 & Sharedvideosotherssentyousavedonyourdevice & 30.56\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosotherssentyousavedonyourdeviceAPPSERVER} \\ 
50 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 30.0\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouareveryaccurately(whereyouareinaroom)APPSERVER} \\ 
51 & Learnedhowmuchyouuseyourcomputer & 28.57\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourcomputerAPPSERVER} \\ 
52 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 28.13\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofpeople(withanoutward-facingcamera)atrandomAPPSERVER} \\ 
53 & Sharedphotosyoutookwhicharealreadyonyourdevice & 27.27\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoutookwhicharealreadyonyourdeviceAPPSERVER} \\ 
54 & Learnedwhatproductsyoubuy & 26.32\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatproductsyoubuyAPPSERVER} \\ 
55 & Learnedwhenyouaresickorhealthy & 26.19\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouaresickorhealthyAPPSERVER} \\ 
56 & Trackedwhereyouare(Likeagps) & 25.0\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouare(likeaGPS)APPSERVER} \\ 
57 & Learnedwhenyouwerehappyorhavingfun & 24.32\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfunAPPSERVER} \\ 
58 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 22.58\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)APPSERVER} \\ 
59 & Learnedwhenandhowwellyouaresleeping & 21.43\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowwellyouaresleepingAPPSERVER} \\ 
60 & Copiedanduploadedmusicfromyourdevice & 20.83\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedmusicfromyourdeviceAPPSERVER} \\ 
61 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 18.92\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureAPPSERVER} \\ 
62 & Learnedwhenyouarebusyorinterruptible & 17.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptibleAPPSERVER} \\ 
63 & Learnedwhenhowandhowmuchyouexercise & 16.67\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexerciseAPPSERVER} \\ 
64 & Learnedhowmuchyouuseyourphone & 15.79\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourphoneAPPSERVER} \\ 
65 & Learnedyourage & 14.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourageAPPSERVER} \\ 
66 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 13.79\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowcloseyouaretootherpeopleyouinteractwithAPPSERVER} \\ 
67 & Learnedwhenhowmuchandwhatyouareeating & 12.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowmuchandwhatyouareeatingAPPSERVER} \\ 
68 & Learnedwhattelevisionshowsyouwatch & 11.43\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchAPPSERVER} \\ 
69 & Learnedyourgender & 9.52\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourgenderAPPSERVER} \\ 
70 & Learnedyourheartrate & 9.09\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourheartrateAPPSERVER} \\ 
71 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 6.98\% & \includegraphics[width = 2cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)APPSERVER} \\ 
72 & Learnedthelanguageyouwerespeaking & 2.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedthelanguageyouwerespeakingAPPSERVER} \\ 
\end{longtable}

\end{document}