### 2015-02-26 Tables(2)/collapsed.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR(\%)}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR(\%)}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Tookavideoofyouunclothed & 95.97 & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouunclothedcombined} \\ 
2 & Learnedyourbankaccountinformation & 95.91 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourbankaccountinformationcombined} \\ 
3 & Learnedyoursocialsecuritynumber & 94.84 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyoursocialsecuritynumbercombined} \\ 
4 & Tookavideoofyouenteringinyourpinatanatm & 92.67 & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMcombined} \\ 
5 & Tookaphotoofyouunclothed & 92.59 & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouunclothedcombined} \\ 
6 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 91.39 & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingcombined} \\ 
7 & Learnedyourusernameandpasswordforwebsites & 89.55 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitescombined} \\ 
8 & Learnedyourcreditcardinformation & 88.98 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourcreditcardinformationcombined} \\ 
9 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 88.41 & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingcombined} \\ 
10 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 87.5 & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomecombined} \\ 
 & \vdots & & \\
64 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 40.51 & \includegraphics[width = 2cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\ 
65 & Learnedwhenhowandhowmuchyouexercise & 38.66 & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexercisecombined} \\ 
66 & Learnedwhenyouwerehappyorhavingfun & 34.75 & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfuncombined} \\ 
67 & Learnedwhattelevisionshowsyouwatch & 30.2 & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchcombined} \\ 
68 & Learnedwhenyouarebusyorinterruptible & 29.5 & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptiblecombined} \\ 
69 & Copiedanduploadedmusicfromyourdevice & 28.06 & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedmusicfromyourdevicecombined} \\ 
70 & Learnedyourheartrate & 27.5 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourheartratecombined} \\ 
71 & Learnedyourage & 24.29 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouragecombined} \\ 
72 & Learnedthelanguageyouwerespeaking & 15.86 & \includegraphics[width = 2cm, height = 0.5cm]{learnedthelanguageyouwerespeakingcombined} \\ 
73 & Learnedyourgender & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourgendercombined} \\ 
\end{longtable}

\end{document}