### 2015-02-26 Tables(2)/friends.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Tookavideoofyouenteringinyourpinatanatm & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMFRIENDS} \\  
2 & Learnedyoursocialsecuritynumber & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyoursocialsecuritynumberFRIENDS} \\  
3 & Learnedyourcreditcardinformation & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourcreditcardinformationFRIENDS} \\  
4 & Tookavideoofyouunclothed & 97.83\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouunclothedFRIENDS} \\  
5 & Learnedyourusernameandpasswordforwebsites & 96.77\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitesFRIENDS} \\  
6 & Tookaphotoofyouunclothed & 96.15\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouunclothedFRIENDS} \\  
7 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 95.24\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinadigitalpasscodetoalockeddoorFRIENDS} \\  
8 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 94.87\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingFRIENDS} \\  
9 & Recordedyourworkconversations & 94.74\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourworkconversationsFRIENDS} \\  
10 & Learnedyourbankaccountinformation & 94.44\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourbankaccountinformationFRIENDS} \\  
11 & Recordedyourphoneconversations & 93.55\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourphoneconversationsFRIENDS} \\  
12 & Learnedwhatmedicalconditionsyouhave & 92.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicalconditionsyouhaveFRIENDS} \\  
13 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 91.89\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingFRIENDS} \\  
14 & Learnedhowmuchmoneyyouhave & 90.7\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchmoneyyouhaveFRIENDS} \\  
15 & Learnedyourusernameforwebsites & 90.32\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameforwebsitesFRIENDS} \\  
16 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 87.8\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)athomeFRIENDS} \\  
17 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 87.18\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedaudiorecordingsyoumadeonyourdeviceFRIENDS} \\  
18 & Recordedyourpassingconversations & 86.11\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourpassingconversationsFRIENDS} \\  
19 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomeFRIENDS} \\  
20 & Learnedhowmuchdebtyouhave & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchdebtyouhaveFRIENDS} \\  
21 & Sharedphotosotherssenttoyousavedonyourdevice & 85.19\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosotherssenttoyousavedonyourdeviceFRIENDS} \\  
22 & Tookavideoofyouoff-Guard & 85.19\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouoff-guardFRIENDS} \\  
23 & Tookaphotoofyouoff-Guard & 83.33\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouoff-guardFRIENDS} \\  
24 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 82.86\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)atrandomFRIENDS} \\  
25 & Sharedvideosotherssentyousavedonyourdevice & 82.86\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosotherssentyousavedonyourdeviceFRIENDS} \\  
26 & Tookavideoofyouintoxicated & 81.48\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouintoxicatedFRIENDS} \\  
27 & Tookaphotoofyouintoxicated & 80.43\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouintoxicatedFRIENDS} \\  
28 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyoutalkingtoyourself(makingvoicenotes)FRIENDS} \\  
29 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceFRIENDS} \\  
30 & Learnedwhatmedicationyoutake & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicationyoutakeFRIENDS} \\  
31 & Learnedwhenandhowmuchyouhavesex & 78.72\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouhavesexFRIENDS} \\  
32 & Learnedyourfingerprintsomehow & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourfingerprintsomehowFRIENDS} \\  
33 & Tookphotosatwork(Withanoutward-Facingcamera) & 76.32\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosatwork(withanoutward-facingcamera)FRIENDS} \\  
34 & Sharedphotosyoutookwhicharealreadyonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoutookwhicharealreadyonyourdeviceFRIENDS} \\  
35 & Learnedwhenyouwerelyingnervousorstressed & 74.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerelyingnervousorstressedFRIENDS} \\  
36 & Tookvideosatwork(Withanoutward-Facingcamera) & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosatwork(withanoutward-facingcamera)FRIENDS} \\  
37 & Learnedwhereyouarecurrentlygoing & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouarecurrentlygoingFRIENDS} \\  
38 & Learnedwhatwebsitesyougoto & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatwebsitesyougotoFRIENDS} \\  
39 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 73.53\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouareveryaccurately(whereyouareinaroom)FRIENDS} \\  
40 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 73.17\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)atrandomFRIENDS} \\  
41 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 72.73\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofpeople(withanoutward-facingcamera)atrandomFRIENDS} \\  
42 & Learnedwhoyouwerespendingtimewith & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhoyouwerespendingtimewithFRIENDS} \\  
43 & Recordedthesoundaroundyou & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedthesoundaroundyouFRIENDS} \\  
44 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 70.0\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)FRIENDS} \\  
45 & Learnedwhereyouworksomehow & 69.57\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouworksomehowFRIENDS} \\  
46 & Learnedhowmuchyouuseyourphone & 67.44\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourphoneFRIENDS} \\  
47 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 63.89\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoudownloadedontheinternetsavedonyourdeviceFRIENDS} \\  
48 & Tookapictureofyoursignature & 63.64\% & \includegraphics[width = 2cm, height = 0.5cm]{tookapictureofyoursignatureFRIENDS} \\  
49 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 63.27\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofpeople(withanoutward-facingcamera)atrandomFRIENDS} \\  
50 & Learnedyouraddress & 62.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouraddressFRIENDS} \\  
51 & Learnedwhereyoulivesomehow & 61.54\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyoulivesomehowFRIENDS} \\  
52 & Sharedvideosyoutookwhicharealreadyonyourdevice & 58.82\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoutookwhicharealreadyonyourdeviceFRIENDS} \\  
53 & Trackedwhereyouare(Likeagps) & 57.69\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouare(likeaGPS)FRIENDS} \\  
54 & Learnedwhatproductsyoubuy & 57.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatproductsyoubuyFRIENDS} \\  
55 & Learnedwhenandhowmuchyouspendtimealone & 51.52\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouspendtimealoneFRIENDS} \\  
56 & Learnedhowmuchyouuseyourcomputer & 51.22\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourcomputerFRIENDS} \\  
57 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureFRIENDS} \\  
58 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowcloseyouaretootherpeopleyouinteractwithFRIENDS} \\  
59 & Learnedyourname & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyournameFRIENDS} \\  
60 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 48.84\% & \includegraphics[width = 2cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)FRIENDS} \\  
61 & Learnedwhenandhowwellyouaresleeping & 43.9\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowwellyouaresleepingFRIENDS} \\  
62 & Learnedwhenhowmuchandwhatyouareeating & 41.86\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowmuchandwhatyouareeatingFRIENDS} \\  
63 & Learnedwhenyouarebusyorinterruptible & 40.63\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptibleFRIENDS} \\  
64 & Learnedwhenyouaresickorhealthy & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouaresickorhealthyFRIENDS} \\  
65 & Learnedwhenyouwerehappyorhavingfun & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfunFRIENDS} \\  
66 & Learnedwhattelevisionshowsyouwatch & 38.46\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchFRIENDS} \\  
67 & Learnedwhenhowandhowmuchyouexercise & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexerciseFRIENDS} \\  
68 & Learnedyourheartrate & 21.82\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourheartrateFRIENDS} \\  
69 & Learnedyourage & 17.86\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourageFRIENDS} \\  
70 & Learnedthelanguageyouwerespeaking & 17.14\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedthelanguageyouwerespeakingFRIENDS} \\  
71 & Learnedyourgender & 15.91\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourgenderFRIENDS} \\  
72 & Copiedanduploadedmusicfromyourdevice & 4.76\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedmusicfromyourdeviceFRIENDS} \\  
\end{longtable}

\end{document}