### 2015-02-26 Tables(2)/recipient.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| r | p{3cm} | p{1cm} | p{1cm} | p{2cm} |}
\hline
Rank & Recipient & VUR(\%) & Variance & Distribution\\
\hline
1& Work & 75.16 & 0.7742& \includegraphics[width = 2cm, height = 2cm]{../recipient_work}\\ 
2& Public & 72.41 & 0.9278& \includegraphics[width = 2cm, height = 2cm]{../recipient_pub}\\ 
3& Friends & 69.47 & 1.0764 & \includegraphics[width = 2cm, height = 2cm]{../recipient_friends}\\ 
4& All & 64.78 & 1.2334& \includegraphics[width = 2cm, height = 2cm]{../recipient_total}\\ 
5& Appserver & 42.28 & 1.7516& \includegraphics[width = 2cm, height = 2cm]{../recipient_app}\\ 
\hline
\end{tabular}
\end{center}
\end{table}