### 2015-02-26 Tables(2)/untitled folder 2/friends.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{0.5cm} | p{7cm} | p{1cm} | c |}
\hline
 Rank & Q &  VUR & Distribution  \\ 
\hline
1 & Learnedwhatmedicalconditionsyouhave & 92.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicalconditionsyouhaveFRIENDS} \\ \hline 
2 & Learnedyouraddress & 62.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyouraddressFRIENDS} \\ \hline 
3 & Learnedwhenhowmuchandwhatyouareeating & 41.86\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowmuchandwhatyouareeatingFRIENDS} \\ \hline 
4 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 87.8\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)athomeFRIENDS} \\ \hline 
5 & Tookphotosatwork(Withanoutward-Facingcamera) & 76.32\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosatwork(withanoutward-facingcamera)FRIENDS} \\ \hline 
6 & Sharedvideosyoutookwhicharealreadyonyourdevice & 58.82\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoutookwhicharealreadyonyourdeviceFRIENDS} \\ \hline 
7 & Sharedphotosotherssenttoyousavedonyourdevice & 85.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosotherssenttoyousavedonyourdeviceFRIENDS} \\ \hline 
8 & Tookavideoofyouoff-Guard & 85.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouoff-guardFRIENDS} \\ \hline 
9 & Copiedanduploadedmusicfromyourdevice & 4.76\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedmusicfromyourdeviceFRIENDS} \\ \hline 
10 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)athomeFRIENDS} \\ \hline 
11 & Learnedhowmuchdebtyouhave & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchdebtyouhaveFRIENDS} \\ \hline 
12 & Learnedyourfingerprintsomehowwith & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourfingerprintsomehowwithFRIENDS} \\ \hline 
13 & Tookvideosatwork(Withanoutward-Facingcamera) & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosatwork(withanoutward-facingcamera)FRIENDS} \\ \hline 
14 & Learnedwhereyouarecurrentlygoing & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouarecurrentlygoingFRIENDS} \\ \hline 
15 & Learnedwhatwebsitesyougoto & 74.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatwebsitesyougotoFRIENDS} \\ \hline 
16 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 73.53\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouareveryaccurately(whereyouareinaroom)FRIENDS} \\ \hline 
17 & Learnedwhattelevisionshowsyouwatch & 38.46\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhattelevisionshowsyouwatchFRIENDS} \\ \hline 
18 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceFRIENDS} \\ \hline 
19 & Learnedwhatmedicationyoutake & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicationyoutakeFRIENDS} \\ \hline 
20 & Learnedwhereyouworksomehow & 69.57\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouworksomehowFRIENDS} \\ \hline 
21 & Learnedyourusernameandpasswordforwebsites & 96.77\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameandpasswordforwebsitesFRIENDS} \\ \hline 
22 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 63.27\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofpeople(withanoutward-facingcamera)atrandomFRIENDS} \\ \hline 
23 & Recordedyourworkconversations & 94.74\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourworkconversationsFRIENDS} \\ \hline 
24 & Learnedhowmuchyouuseyourphone & 67.44\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourphoneFRIENDS} \\ \hline 
25 & Learnedthelanguageyouwerespeaking & 17.14\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedthelanguageyouwerespeakingFRIENDS} \\ \hline 
26 & Tookavideoofyouintoxicated & 81.48\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouintoxicatedFRIENDS} \\ \hline 
27 & Trackedwhereyouare(Likeagps) & 57.69\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouare(likeaGPS)FRIENDS} \\ \hline 
28 & Learnedwhenandhowmuchyouspendtimealone & 51.52\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouspendtimealoneFRIENDS} \\ \hline 
29 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureFRIENDS} \\ \hline 
30 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowcloseyouaretootherpeopleyouinteractwithFRIENDS} \\ \hline 
31 & Learnedyourname & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyournameFRIENDS} \\ \hline 
32 & Learnedhowmuchmoneyyouhave & 90.7\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchmoneyyouhaveFRIENDS} \\ \hline 
33 & Learnedwhoyouwerespendingtimewith & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhoyouwerespendingtimewithFRIENDS} \\ \hline 
34 & Recordedthesoundaroundyou & 71.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedthesoundaroundyouFRIENDS} \\ \hline 
35 & Learnedyourusernameforwebsites & 90.32\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameforwebsitesFRIENDS} \\ \hline 
36 & Tookavideoofyouunclothed & 97.83\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouunclothedFRIENDS} \\ \hline 
37 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 91.89\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingvideoofyoudoingsomethingembarrassingFRIENDS} \\ \hline 
38 & Learnedhowmuchyouuseyourcomputer & 51.22\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourcomputerFRIENDS} \\ \hline 
39 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 82.86\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)atrandomFRIENDS} \\ \hline 
40 & Sharedvideosotherssentyousavedonyourdevice & 82.86\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosotherssentyousavedonyourdeviceFRIENDS} \\ \hline 
41 & Learnedwhenhowandhowmuchyouexercise & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowandhowmuchyouexerciseFRIENDS} \\ \hline 
42 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 70.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)FRIENDS} \\ \hline 
43 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 72.73\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofpeople(withanoutward-facingcamera)atrandomFRIENDS} \\ \hline 
44 & Learnedwhenandhowwellyouaresleeping & 43.9\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowwellyouaresleepingFRIENDS} \\ \hline 
45 & Sharedphotosyoutookwhicharealreadyonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoutookwhicharealreadyonyourdeviceFRIENDS} \\ \hline 
46 & Recordedyourphoneconversations & 93.55\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourphoneconversationsFRIENDS} \\ \hline 
47 & Tookapictureofyoursignature & 63.64\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookapictureofyoursignatureFRIENDS} \\ \hline 
48 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyoutalkingtoyourself(makingvoicenotes)FRIENDS} \\ \hline 
49 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 94.87\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingphotoofyoudoingsomethingembarrassingFRIENDS} \\ \hline 
50 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 95.24\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinadigitalpasscodetoalockeddoorFRIENDS} \\ \hline 
51 & Learnedyourage & 17.86\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourageFRIENDS} \\ \hline 
52 & Learnedwhenyouarebusyorinterruptible & 40.63\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouarebusyorinterruptibleFRIENDS} \\ \hline 
53 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 48.84\% & \includegraphics[width = 2cm, height = 0.5cm]{../scannedyoureyetolearnyoureyepatterns(foreyetracking)FRIENDS} \\ \hline 
54 & Learnedyourgender & 15.91\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourgenderFRIENDS} \\ \hline 
55 & Learnedyourbankaccountinformation & 94.44\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourbankaccountinformationFRIENDS} \\ \hline 
56 & Tookaphotoofyouoff-Guard & 83.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouoff-guardFRIENDS} \\ \hline 
57 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 73.17\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)atrandomFRIENDS} \\ \hline 
58 & Tookavideoofyouenteringinyourpinatanatm & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinyourPINatanATMFRIENDS} \\ \hline 
59 & Learnedyoursocialsecuritynumber & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyoursocialsecuritynumberFRIENDS} \\ \hline 
60 & Learnedyourcreditcardinformation & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourcreditcardinformationFRIENDS} \\ \hline 
61 & Tookaphotoofyouintoxicated & 80.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouintoxicatedFRIENDS} \\ \hline 
62 & Tookaphotoofyouunclothed & 96.15\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouunclothedFRIENDS} \\ \hline 
63 & Learnedwhenandhowmuchyouhavesex & 78.72\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouhavesexFRIENDS} \\ \hline 
64 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 63.89\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoudownloadedontheinternetsavedonyourdeviceFRIENDS} \\ \hline 
65 & Learnedwhenyouwerelyingnervousorstressed & 74.29\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerelyingnervousorstressedFRIENDS} \\ \hline 
66 & Learnedwhenyouaresickorhealthy & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouaresickorhealthyFRIENDS} \\ \hline 
67 & Learnedwhenyouwerehappyorhavingfun & 40.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerehappyorhavingfunFRIENDS} \\ \hline 
68 & Recordedyourpassingconversations & 86.11\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourpassingconversationsFRIENDS} \\ \hline 
69 & Learnedyourheartrate & 21.82\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourheartrateFRIENDS} \\ \hline 
70 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 87.18\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedaudiorecordingsyoumadeonyourdeviceFRIENDS} \\ \hline 
71 & Learnedwhereyoulivesomehow & 61.54\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyoulivesomehowFRIENDS} \\ \hline 
72 & Learnedwhatproductsyoubuy & 57.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatproductsyoubuyFRIENDS} \\ \hline 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}