### 2015-02-26 Tables(2)/work.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Recordedyourworkconversations & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourworkconversationsWORKCONTACTS} \\  
2 & Tookaphotoofyouunclothed & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouunclothedWORKCONTACTS} \\  
3 & Tookavideoofyouunclothed & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouunclothedWORKCONTACTS} \\  
4 & Learnedyourbankaccountinformation & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourbankaccountinformationWORKCONTACTS} \\  
5 & Learnedyourcreditcardinformation & 95.12\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourcreditcardinformationWORKCONTACTS} \\  
6 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 94.59\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingWORKCONTACTS} \\  
7 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 94.29\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingWORKCONTACTS} \\  
8 & Learnedyourusernameandpasswordforwebsites & 94.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitesWORKCONTACTS} \\  
9 & Learnedyoursocialsecuritynumber & 93.75\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyoursocialsecuritynumberWORKCONTACTS} \\  
10 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 93.55\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomeWORKCONTACTS} \\  
11 & Learnedhowmuchmoneyyouhave & 93.1\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchmoneyyouhaveWORKCONTACTS} \\  
12 & Learnedwhenyouwerelyingnervousorstressed & 91.89\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerelyingnervousorstressedWORKCONTACTS} \\  
13 & Tookavideoofyouoff-Guard & 91.43\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouoff-guardWORKCONTACTS} \\  
14 & Recordedyourphoneconversations & 90.32\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourphoneconversationsWORKCONTACTS} \\  
15 & Learnedwhenandhowmuchyouhavesex & 90.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouhavesexWORKCONTACTS} \\  
16 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 89.19\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)athomeWORKCONTACTS} \\  
17 & Tookavideoofyouintoxicated & 88.89\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouintoxicatedWORKCONTACTS} \\  
18 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 88.57\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyoutalkingtoyourself(makingvoicenotes)WORKCONTACTS} \\  
19 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 88.24\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)atrandomWORKCONTACTS} \\  
20 & Recordedyourpassingconversations & 87.5\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourpassingconversationsWORKCONTACTS} \\  
21 & Tookavideoofyouenteringinyourpinatanatm & 87.5\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMWORKCONTACTS} \\  
22 & Learnedhowmuchdebtyouhave & 87.18\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchdebtyouhaveWORKCONTACTS} \\  
23 & Tookaphotoofyouintoxicated & 87.1\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouintoxicatedWORKCONTACTS} \\  
24 & Learnedwhatwebsitesyougoto & 86.84\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatwebsitesyougotoWORKCONTACTS} \\  
25 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 86.27\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoudownloadedfromtheinternetsavedonyourdeviceWORKCONTACTS} \\  
26 & Tookapictureofyoursignature & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{tookapictureofyoursignatureWORKCONTACTS} \\  
27 & Learnedwhatmedicalconditionsyouhave & 85.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicalconditionsyouhaveWORKCONTACTS} \\  
28 & Learnedyourusernameforwebsites & 84.62\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameforwebsitesWORKCONTACTS} \\  
29 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 82.86\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoudownloadedontheinternetsavedonyourdeviceWORKCONTACTS} \\  
30 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 82.05\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofpeople(withanoutward-facingcamera)atrandomWORKCONTACTS} \\  
31 & Learnedyouraddress & 81.58\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouraddressWORKCONTACTS} \\  
32 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 81.48\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)atrandomWORKCONTACTS} \\  
33 & Tookphotosatwork(Withanoutward-Facingcamera) & 81.48\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosatwork(withanoutward-facingcamera)WORKCONTACTS} \\  
34 & Learnedwhatmedicationyoutake & 81.25\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicationyoutakeWORKCONTACTS} \\  
35 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 81.08\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinadigitalpasscodetoalockeddoorWORKCONTACTS} \\  
36 & Learnedwhenandhowmuchyouspendtimealone & 80.95\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouspendtimealoneWORKCONTACTS} \\  
37 & Tookaphotoofyouoff-Guard & 80.49\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouoff-guardWORKCONTACTS} \\  
38 & Sharedvideosotherssentyousavedonyourdevice & 80.49\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosotherssentyousavedonyourdeviceWORKCONTACTS} \\  
39 & Sharedphotosyoutookwhicharealreadyonyourdevice & 79.55\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoutookwhicharealreadyonyourdeviceWORKCONTACTS} \\  
40 & Sharedvideosyoutookwhicharealreadyonyourdevice & 79.31\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoutookwhicharealreadyonyourdeviceWORKCONTACTS} \\  
41 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 78.13\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouareveryaccurately(whereyouareinaroom)WORKCONTACTS} \\  
42 & Learnedwhoyouwerespendingtimewith & 76.19\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhoyouwerespendingtimewithWORKCONTACTS} \\  
43 & Recordedthesoundaroundyou & 75.76\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedthesoundaroundyouWORKCONTACTS} \\  
44 & Sharedphotosotherssenttoyousavedonyourdevice & 75.61\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosotherssenttoyousavedonyourdeviceWORKCONTACTS} \\  
45 & Learnedwhenhowmuchandwhatyouareeating & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowmuchandwhatyouareeatingWORKCONTACTS} \\  
46 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.22\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedaudiorecordingsyoumadeonyourdeviceWORKCONTACTS} \\  
47 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 71.88\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofpeople(withanoutward-facingcamera)atrandomWORKCONTACTS} \\  
48 & Learnedyourfingerprintsomehow & 70.27\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourfingerprintsomehowWORKCONTACTS} \\  
49 & Tookvideosatwork(Withanoutward-Facingcamera) & 70.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosatwork(withanoutward-facingcamera)WORKCONTACTS} \\  
50 & Learnedwhereyoulivesomehow & 69.44\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyoulivesomehowWORKCONTACTS} \\  
51 & Learnedhowmuchyouuseyourphone & 68.75\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourphoneWORKCONTACTS} \\  
52 & Learnedwhenhowandhowmuchyouexercise & 66.67\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexerciseWORKCONTACTS} \\  
53 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 66.67\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)WORKCONTACTS} \\  
54 & Learnedwhereyouarecurrentlygoing & 65.85\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouarecurrentlygoingWORKCONTACTS} \\  
55 & Trackedwhereyouare(Likeagps) & 63.64\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouare(likeaGPS)WORKCONTACTS} \\  
56 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 63.16\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowyouwerefeelingbasedonheartratebreathingandortemperatureWORKCONTACTS} \\  
57 & Learnedwhatproductsyoubuy & 62.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatproductsyoubuyWORKCONTACTS} \\  
58 & Learnedwhenyouaresickorhealthy & 62.22\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouaresickorhealthyWORKCONTACTS} \\  
59 & Learnedwhenandhowwellyouaresleeping & 62.07\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowwellyouaresleepingWORKCONTACTS} \\  
60 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 61.11\% & \includegraphics[width = 2cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)WORKCONTACTS} \\  
61 & Learnedwhereyouworksomehow & 57.14\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouworksomehowWORKCONTACTS} \\  
62 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 51.35\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowcloseyouaretootherpeopleyouinteractwithWORKCONTACTS} \\  
63 & Learnedhowmuchyouuseyourcomputer & 45.45\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourcomputerWORKCONTACTS} \\  
64 & Learnedyourheartrate & 44.12\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourheartrateWORKCONTACTS} \\  
65 & Learnedwhenyouwerehappyorhavingfun & 43.75\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfunWORKCONTACTS} \\  
66 & Copiedanduploadedmusicfromyourdevice & 42.11\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedmusicfromyourdeviceWORKCONTACTS} \\  
67 & Learnedyourage & 36.67\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourageWORKCONTACTS} \\  
68 & Learnedwhenyouarebusyorinterruptible & 36.36\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptibleWORKCONTACTS} \\  
69 & Learnedwhattelevisionshowsyouwatch & 36.36\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchWORKCONTACTS} \\  
70 & Learnedthelanguageyouwerespeaking & 28.13\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedthelanguageyouwerespeakingWORKCONTACTS} \\  
71 & Learnedyourname & 26.47\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyournameWORKCONTACTS} \\  
72 & Learnedyourgender & 15.79\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourgenderWORKCONTACTS} \\  
\end{longtable}

\end{document}