### 2015-03-03 tables-more/benefit.tex
started to look at results

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{2cm} | p{1cm} | p{1cm} | p{1cm} | c |}
\hline
Technology & Q1 &  Median & Q3 & Distribution  \\ 
\hline
Voice Based Emotion Detectionben & 20.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../voicebasedemotiondetectionben} \\ 
Emailben & 50.0 & 29.0 & 77.5 & \includegraphics[width = 2cm, height = 0.5cm]{../emailben} \\ 
Internetben & 65.0 & 45.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{../internetben} \\ 
Age Detectionben & 12.0 & 10.0 & 22.0 & \includegraphics[width = 2cm, height = 0.5cm]{../agedetectionben} \\ 
Laptopsben & 60.0 & 40.0 & 80.0 & \includegraphics[width = 2cm, height = 0.5cm]{../laptopsben} \\ 
Gender Detectionben & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{../genderdetectionben} \\ 
Facial Detectionben & 20.0 & 10.0 & 34.0 & \includegraphics[width = 2cm, height = 0.5cm]{../facialdetectionben} \\ 
Discreet Video Cameraben & 20.0 & 15.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../discreetvideocameraben} \\ 
Facial Recognitionben & 22.0 & 12.5 & 42.5 & \includegraphics[width = 2cm, height = 0.5cm]{../facialrecognitionben} \\ 
Heart Rate Detectionben & 40.0 & 26.0 & 65.0 & \includegraphics[width = 2cm, height = 0.5cm]{../heartratedetectionben} \\ 
Google Glassben & 20.0 & 12.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../googleglassben} \\ 
Speech To Textben & 25.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../speechtotextben} \\ 
Discreet Microphoneben & 15.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{../discreetmicrophoneben} \\ 
Fitness Trackersben & 18.5 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../fitnesstrackersben} \\ 
Voice Recognitionben & 25.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../voicerecognitionben} \\ 
Smartwatchesben & 20.0 & 10.0 & 35.0 & \includegraphics[width = 2cm, height = 0.5cm]{../smartwatchesben} \\ 
Language Detectionben & 35.0 & 15.0 & 60.0 & \includegraphics[width = 2cm, height = 0.5cm]{../languagedetectionben} \\ 
Motorcycle & 20.0 & 12.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../MotorcycleBenefit} \\ 
Cubetasticben & 15.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../cubetasticben} \\ 
Lawnmower & 24.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../LawnmowerBenefit} \\ 
Electricity & 88.0 & 50.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{../ElectricityBenefit} \\ 
Smartphonesben & 50.0 & 30.0 & 75.0 & \includegraphics[width = 2cm, height = 0.5cm]{../smartphonesben} \\ 
Ben & 100.0 & nan & nan & \includegraphics[width = 2cm, height = 0.5cm]{../ben} \\ 
Handgun & 20.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../HandgunBenefit} \\ 
Location Trackingben & 40.0 & 20.0 & 70.0 & \includegraphics[width = 2cm, height = 0.5cm]{../locationtrackingben} \\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}