### 2015-03-03 tables-more/pub.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingphotoofyoudoingsomethingembarrassingPUBLIC} \\  
2 & Learnedyoursocialsecuritynumber & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyoursocialsecuritynumberPUBLIC} \\  
3 & Learnedhowmuchmoneyyouhave & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchmoneyyouhavePUBLIC} \\  
4 & Recordedyourphoneconversations & 97.22\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourphoneconversationsPUBLIC} \\  
5 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 96.97\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)athomePUBLIC} \\  
6 & Recordedyourworkconversations & 96.67\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourworkconversationsPUBLIC} \\  
7 & Sharedvideosotherssentyousavedonyourdevice & 95.83\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosotherssentyousavedonyourdevicePUBLIC} \\  
8 & Learnedyourbankaccountinformation & 95.83\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourbankaccountinformationPUBLIC} \\  
9 & Learnedyourusernameandpasswordforwebsites & 95.35\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameandpasswordforwebsitesPUBLIC} \\  
10 & Tookavideoofyouunclothed & 94.44\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouunclothedPUBLIC} \\  
11 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 94.12\% & \includegraphics[width = 2cm, height = 0.5cm]{tookanincriminatingvideoofyoudoingsomethingembarrassingPUBLIC} \\  
12 & Tookavideoofyouenteringinyourpinatanatm & 93.94\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinyourPINatanATMPUBLIC} \\  
13 & Learnedyourcreditcardinformation & 93.55\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourcreditcardinformationPUBLIC} \\  
14 & Learnedyouraddress & 93.18\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouraddressPUBLIC} \\  
15 & Trackedwhereyouare(Likeagps) & 92.11\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouare(likeaGPS)PUBLIC} \\  
16 & Tookaphotoofyouunclothed & 91.67\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouunclothedPUBLIC} \\  
17 & Tookavideoofyouintoxicated & 91.67\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouintoxicatedPUBLIC} \\  
18 & Tookaphotoofyouintoxicated & 90.7\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouintoxicatedPUBLIC} \\  
19 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 90.24\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)atrandomPUBLIC} \\  
20 & Learnedhowmuchdebtyouhave & 90.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchdebtyouhavePUBLIC} \\  
21 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 89.19\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouenteringinadigitalpasscodetoalockeddoorPUBLIC} \\  
22 & Learnedwhereyoulivesomehow & 87.88\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyoulivesomehowPUBLIC} \\  
23 & Learnedwhenandhowmuchyouhavesex & 87.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouhavesexPUBLIC} \\  
24 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 86.49\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofyou(withaninward-facingcamera)athomePUBLIC} \\  
25 & Learnedwhatmedicalconditionsyouhave & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicalconditionsyouhavePUBLIC} \\  
26 & Recordedyourpassingconversations & 85.29\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyourpassingconversationsPUBLIC} \\  
27 & Sharedphotosotherssenttoyousavedonyourdevice & 84.62\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosotherssenttoyousavedonyourdevicePUBLIC} \\  
28 & Tookvideosatwork(Withanoutward-Facingcamera) & 83.87\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosatwork(withanoutward-facingcamera)PUBLIC} \\  
29 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 83.33\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofyou(withaninward-facingcamera)atrandomPUBLIC} \\  
30 & Sharedvideosyoutookwhicharealreadyonyourdevice & 82.93\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoutookwhicharealreadyonyourdevicePUBLIC} \\  
31 & Tookphotosatwork(Withanoutward-Facingcamera) & 82.14\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosatwork(withanoutward-facingcamera)PUBLIC} \\  
32 & Learnedyourfingerprintsomehow & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourfingerprintsomehowPUBLIC} \\  
33 & Tookaphotoofyouoff-Guard & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{tookaphotoofyouoff-guardPUBLIC} \\  
34 & Tookavideoofyouoff-Guard & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{tookavideoofyouoff-guardPUBLIC} \\  
35 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 78.05\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedyoutalkingtoyourself(makingvoicenotes)PUBLIC} \\  
36 & Sharedphotosyoutookwhicharealreadyonyourdevice & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoutookwhicharealreadyonyourdevicePUBLIC} \\  
37 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 76.19\% & \includegraphics[width = 2cm, height = 0.5cm]{trackedwhereyouareveryaccurately(whereyouareinaroom)PUBLIC} \\  
38 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 76.09\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedphotosyoudownloadedfromtheinternetsavedonyourdevicePUBLIC} \\  
39 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 75.61\% & \includegraphics[width = 2cm, height = 0.5cm]{tookvideosofpeople(withanoutward-facingcamera)atrandomPUBLIC} \\  
40 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedaudiorecordingsyoumadeonyourdevicePUBLIC} \\  
41 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{sharedvideosyoudownloadedontheinternetsavedonyourdevicePUBLIC} \\  
42 & Learnedwhenyouwerelyingnervousorstressed & 74.36\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerelyingnervousorstressedPUBLIC} \\  
43 & Learnedyourusernameforwebsites & 74.29\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourusernameforwebsitesPUBLIC} \\  
44 & Learnedwhatmedicationyoutake & 73.68\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatmedicationyoutakePUBLIC} \\  
45 & Learnedwhatwebsitesyougoto & 70.27\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatwebsitesyougotoPUBLIC} \\  
46 & Learnedwhenandhowmuchyouspendtimealone & 69.44\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowmuchyouspendtimealonePUBLIC} \\  
47 & Learnedyourname & 68.97\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyournamePUBLIC} \\  
48 & Learnedwhereyouarecurrentlygoing & 68.75\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouarecurrentlygoingPUBLIC} \\  
49 & Learnedwhereyouworksomehow & 64.86\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhereyouworksomehowPUBLIC} \\  
50 & Recordedthesoundaroundyou & 64.71\% & \includegraphics[width = 2cm, height = 0.5cm]{recordedthesoundaroundyouPUBLIC} \\  
51 & Tookapictureofyoursignature & 64.52\% & \includegraphics[width = 2cm, height = 0.5cm]{tookapictureofyoursignaturePUBLIC} \\  
52 & Learnedwhenyouaresickorhealthy & 61.9\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouaresickorhealthyPUBLIC} \\  
53 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 61.22\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowcloseyouaretootherpeopleyouinteractwithPUBLIC} \\  
54 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 61.11\% & \includegraphics[width = 2cm, height = 0.5cm]{tookphotosofpeople(withanoutward-facingcamera)atrandomPUBLIC} \\  
55 & Learnedwhoyouwerespendingtimewith & 60.0\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhoyouwerespendingtimewithPUBLIC} \\  
56 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 59.46\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)PUBLIC} \\  
57 & Learnedhowmuchyouuseyourphone & 56.1\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourphonePUBLIC} \\  
58 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 55.56\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturePUBLIC} \\  
59 & Learnedwhatproductsyoubuy & 55.56\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhatproductsyoubuyPUBLIC} \\  
60 & Learnedhowmuchyouuseyourcomputer & 52.5\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedhowmuchyouuseyourcomputerPUBLIC} \\  
61 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{scannedyoureyetolearnyoureyepatterns(foreyetracking)PUBLIC} \\  
62 & Learnedwhenhowmuchandwhatyouareeating & 45.71\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowmuchandwhatyouareeatingPUBLIC} \\  
63 & Learnedwhenandhowwellyouaresleeping & 41.94\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenandhowwellyouaresleepingPUBLIC} \\  
64 & Copiedanduploadedmusicfromyourdevice & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{copiedanduploadedmusicfromyourdevicePUBLIC} \\  
65 & Learnedyourheartrate & 36.84\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourheartratePUBLIC} \\  
66 & Learnedwhenhowandhowmuchyouexercise & 34.48\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenhowandhowmuchyouexercisePUBLIC} \\  
67 & Learnedyourage & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyouragePUBLIC} \\  
68 & Learnedwhattelevisionshowsyouwatch & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhattelevisionshowsyouwatchPUBLIC} \\  
69 & Learnedwhenyouwerehappyorhavingfun & 32.43\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouwerehappyorhavingfunPUBLIC} \\  
70 & Learnedwhenyouarebusyorinterruptible & 28.89\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedwhenyouarebusyorinterruptiblePUBLIC} \\  
71 & Learnedyourgender & 19.44\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedyourgenderPUBLIC} \\  
72 & Learnedthelanguageyouwerespeaking & 18.42\% & \includegraphics[width = 2cm, height = 0.5cm]{learnedthelanguageyouwerespeakingPUBLIC} \\  
\end{longtable}

\end{document}