### 2015-03-03 tables-more/untitled folder 2/benefit.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{2cm} | p{1cm} | p{1cm} | p{1cm} | c |}
\hline
Technology & Q1 &  Median & Q3 & Distribution  \\ 
\hline
Language Detection  & 15.0 & 35.0 & 60.0 & \includegraphics[width = 2cm, height = 0.5cm]{../languagedetectionben} \\ 
Email  & 29.0 & 50.0 & 77.5 & \includegraphics[width = 2cm, height = 0.5cm]{../emailben} \\ 
Heart Rate Detection  & 26.0 & 40.0 & 65.0 & \includegraphics[width = 2cm, height = 0.5cm]{../heartratedetectionben} \\ 
Cubetastic  & 10.0 & 15.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../cubetasticben} \\ 
Speech To Text  & 15.0 & 25.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../speechtotextben} \\ 
Location Tracking  & 20.0 & 40.0 & 70.0 & \includegraphics[width = 2cm, height = 0.5cm]{../locationtrackingben} \\ 
Facial Detection  & 10.0 & 20.0 & 34.0 & \includegraphics[width = 2cm, height = 0.5cm]{../facialdetectionben} \\ 
Age Detection  & 10.0 & 12.0 & 22.0 & \includegraphics[width = 2cm, height = 0.5cm]{../agedetectionben} \\ 
Discreet Video Camera  & 15.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../discreetvideocameraben} \\ 
Facial Recognition  & 12.5 & 22.0 & 42.5 & \includegraphics[width = 2cm, height = 0.5cm]{../facialrecognitionben} \\ 
Internet  & 45.0 & 65.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{../internetben} \\ 
Voice Recognition  & 15.0 & 25.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../voicerecognitionben} \\ 
Discreet Microphone  & 10.0 & 15.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{../discreetmicrophoneben} \\ 
Motorcycle & 12.0 & 20.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../MotorcycleBenefit} \\ 
Gender Detection  & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{../genderdetectionben} \\ 
Voice Based Emotion Detection  & 10.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../voicebasedemotiondetectionben} \\ 
Smartwatches  & 10.0 & 20.0 & 35.0 & \includegraphics[width = 2cm, height = 0.5cm]{../smartwatchesben} \\ 
Electricity & 50.0 & 88.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{../ElectricityBenefit} \\ 
Google Glass  & 12.0 & 20.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../googleglassben} \\ 
Laptops  & 40.0 & 60.0 & 80.0 & \includegraphics[width = 2cm, height = 0.5cm]{../laptopsben} \\ 
Lawnmower & 15.0 & 24.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{../LawnmowerBenefit} \\ 
Handgun & 10.0 & 20.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../HandgunBenefit} \\ 
Smartphones  & 30.0 & 50.0 & 75.0 & \includegraphics[width = 2cm, height = 0.5cm]{../smartphonesben} \\ 
Fitness Trackers  & 10.0 & 18.5 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{../fitnesstrackersben} \\ 
\hline
\end{tabular}
\caption{Benefit Table}
\label{top10}
\end{center}
\end{table}