### 2015-03-03 tables-more/untitled folder 2/pub.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{0.5cm} | p{7cm} | p{1cm} | c |}
\hline
 Rank & Q &  VUR & Distribution  \\ 
\hline
1 & Learnedwhenandhowmuchyouhavesex & 87.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouhavesexPUBLIC} \\ \hline 
2 & Copiedanduploadedmusicfromyourdevice & 37.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedmusicfromyourdevicePUBLIC} \\ \hline 
3 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 61.11\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofpeople(withanoutward-facingcamera)atrandomPUBLIC} \\ \hline 
4 & Learnedyourcreditcardinformation & 93.55\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourcreditcardinformationPUBLIC} \\ \hline 
5 & Learnedwhenandhowmuchyouspendtimealone & 69.44\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowmuchyouspendtimealonePUBLIC} \\ \hline 
6 & Learnedwhereyouarecurrentlygoing & 68.75\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouarecurrentlygoingPUBLIC} \\ \hline 
7 & Tookphotosatwork(Withanoutward-Facingcamera) & 82.14\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosatwork(withanoutward-facingcamera)PUBLIC} \\ \hline 
8 & Tookavideoofyouunclothed & 94.44\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouunclothedPUBLIC} \\ \hline 
9 & Learnedwhenhowandhowmuchyouexercise & 34.48\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowandhowmuchyouexercisePUBLIC} \\ \hline 
10 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 94.12\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingvideoofyoudoingsomethingembarrassingPUBLIC} \\ \hline 
11 & Sharedphotosyoutookwhicharealreadyonyourdevice & 77.78\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoutookwhicharealreadyonyourdevicePUBLIC} \\ \hline 
12 & Trackedwhereyouare(Likeagps) & 92.11\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouare(likeaGPS)PUBLIC} \\ \hline 
13 & Tookavideoofyouenteringinyourpinatanatm & 93.94\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinyourPINatanATMPUBLIC} \\ \hline 
14 & Learnedyourusernameandpasswordforwebsites & 95.35\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameandpasswordforwebsitesPUBLIC} \\ \hline 
15 & Learnedwhatwebsitesyougoto & 70.27\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatwebsitesyougotoPUBLIC} \\ \hline 
16 & Recordedthesoundaroundyou & 64.71\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedthesoundaroundyouPUBLIC} \\ \hline 
17 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 59.46\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)PUBLIC} \\ \hline 
18 & Learnedyourusernameforwebsites & 74.29\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourusernameforwebsitesPUBLIC} \\ \hline 
19 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 90.24\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)atrandomPUBLIC} \\ \hline 
20 & Learnedwhatmedicationyoutake & 73.68\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicationyoutakePUBLIC} \\ \hline 
21 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 89.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouenteringinadigitalpasscodetoalockeddoorPUBLIC} \\ \hline 
22 & Learnedwhereyoulivesomehow & 87.88\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyoulivesomehowPUBLIC} \\ \hline 
23 & Learnedwhenyouarebusyorinterruptible & 28.89\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouarebusyorinterruptiblePUBLIC} \\ \hline 
24 & Learnedyouraddress & 93.18\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyouraddressPUBLIC} \\ \hline 
25 & Learnedwhenandhowwellyouaresleeping & 41.94\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenandhowwellyouaresleepingPUBLIC} \\ \hline 
26 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 50.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../scannedyoureyetolearnyoureyepatterns(foreyetracking)PUBLIC} \\ \hline 
27 & Tookaphotoofyouintoxicated & 90.7\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouintoxicatedPUBLIC} \\ \hline 
28 & Learnedwhenhowmuchandwhatyouareeating & 45.71\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenhowmuchandwhatyouareeatingPUBLIC} \\ \hline 
29 & Tookapictureofyoursignature & 64.52\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookapictureofyoursignaturePUBLIC} \\ \hline 
30 & Learnedyourname & 68.97\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyournamePUBLIC} \\ \hline 
31 & Learnedwhereyouworksomehow & 64.86\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhereyouworksomehowPUBLIC} \\ \hline 
32 & Learnedwhoyouwerespendingtimewith & 60.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhoyouwerespendingtimewithPUBLIC} \\ \hline 
33 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 76.09\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosyoudownloadedfromtheinternetsavedonyourdevicePUBLIC} \\ \hline 
34 & Learnedwhenyouwerehappyorhavingfun & 32.43\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerehappyorhavingfunPUBLIC} \\ \hline 
35 & Learnedyourage & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyouragePUBLIC} \\ \hline 
36 & Learnedwhattelevisionshowsyouwatch & 33.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhattelevisionshowsyouwatchPUBLIC} \\ \hline 
37 & Tookvideosatwork(Withanoutward-Facingcamera) & 83.87\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosatwork(withanoutward-facingcamera)PUBLIC} \\ \hline 
38 & Sharedvideosotherssentyousavedonyourdevice & 95.83\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosotherssentyousavedonyourdevicePUBLIC} \\ \hline 
39 & Learnedyourbankaccountinformation & 95.83\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourbankaccountinformationPUBLIC} \\ \hline 
40 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 78.05\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyoutalkingtoyourself(makingvoicenotes)PUBLIC} \\ \hline 
41 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../copiedanduploadedaudiorecordingsyoumadeonyourdevicePUBLIC} \\ \hline 
42 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 75.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoudownloadedontheinternetsavedonyourdevicePUBLIC} \\ \hline 
43 & Learnedwhenyouaresickorhealthy & 61.9\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouaresickorhealthyPUBLIC} \\ \hline 
44 & Learnedyourfingerprintsomehow & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourfingerprintsomehowPUBLIC} \\ \hline 
45 & Tookaphotoofyouoff-Guard & 80.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouoff-guardPUBLIC} \\ \hline 
46 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 55.56\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturePUBLIC} \\ \hline 
47 & Learnedwhatproductsyoubuy & 55.56\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatproductsyoubuyPUBLIC} \\ \hline 
48 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 86.49\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofyou(withaninward-facingcamera)athomePUBLIC} \\ \hline 
49 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 61.22\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowcloseyouaretootherpeopleyouinteractwithPUBLIC} \\ \hline 
50 & Tookavideoofyouoff-Guard & 79.17\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouoff-guardPUBLIC} \\ \hline 
51 & Learnedhowmuchdebtyouhave & 90.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchdebtyouhavePUBLIC} \\ \hline 
52 & Learnedwhatmedicalconditionsyouhave & 85.71\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhatmedicalconditionsyouhavePUBLIC} \\ \hline 
53 & Learnedyourgender & 19.44\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourgenderPUBLIC} \\ \hline 
54 & Learnedthelanguageyouwerespeaking & 18.42\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedthelanguageyouwerespeakingPUBLIC} \\ \hline 
55 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 76.19\% & \includegraphics[width = 2cm, height = 0.5cm]{../trackedwhereyouareveryaccurately(whereyouareinaroom)PUBLIC} \\ \hline 
56 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 83.33\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)atrandomPUBLIC} \\ \hline 
57 & Recordedyourworkconversations & 96.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourworkconversationsPUBLIC} \\ \hline 
58 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookanincriminatingphotoofyoudoingsomethingembarrassingPUBLIC} \\ \hline 
59 & Learnedyoursocialsecuritynumber & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyoursocialsecuritynumberPUBLIC} \\ \hline 
60 & Learnedhowmuchmoneyyouhave & 100.0\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchmoneyyouhavePUBLIC} \\ \hline 
61 & Sharedvideosyoutookwhicharealreadyonyourdevice & 82.93\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedvideosyoutookwhicharealreadyonyourdevicePUBLIC} \\ \hline 
62 & Sharedphotosotherssenttoyousavedonyourdevice & 84.62\% & \includegraphics[width = 2cm, height = 0.5cm]{../sharedphotosotherssenttoyousavedonyourdevicePUBLIC} \\ \hline 
63 & Tookaphotoofyouunclothed & 91.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookaphotoofyouunclothedPUBLIC} \\ \hline 
64 & Tookavideoofyouintoxicated & 91.67\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookavideoofyouintoxicatedPUBLIC} \\ \hline 
65 & Recordedyourphoneconversations & 97.22\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourphoneconversationsPUBLIC} \\ \hline 
66 & Recordedyourpassingconversations & 85.29\% & \includegraphics[width = 2cm, height = 0.5cm]{../recordedyourpassingconversationsPUBLIC} \\ \hline 
67 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 75.61\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookvideosofpeople(withanoutward-facingcamera)atrandomPUBLIC} \\ \hline 
68 & Learnedwhenyouwerelyingnervousorstressed & 74.36\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedwhenyouwerelyingnervousorstressedPUBLIC} \\ \hline 
69 & Learnedhowmuchyouuseyourphone & 56.1\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourphonePUBLIC} \\ \hline 
70 & Learnedyourheartrate & 36.84\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedyourheartratePUBLIC} \\ \hline 
71 & Learnedhowmuchyouuseyourcomputer & 52.5\% & \includegraphics[width = 2cm, height = 0.5cm]{../learnedhowmuchyouuseyourcomputerPUBLIC} \\ \hline 
72 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 96.97\% & \includegraphics[width = 2cm, height = 0.5cm]{../tookphotosofyou(withaninward-facingcamera)athomePUBLIC} \\ \hline 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}