### 2015-03-03 collapsed.tex
collapsed/collapsedstd/recipient, and images
title

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} |p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Standard Deviation}} &\multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} & \multicolumn{1}{c|}{\textbf{Standard Deviation}} & \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Tookavideoofyouunclothed & 95.97&0.3115 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouunclothedcombined} \\ 
2 & Learnedyourbankaccountinformation & 95.91&0.3539 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourbankaccountinformationcombined} \\ 
3 & Learnedyoursocialsecuritynumber & 94.84&0.26 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyoursocialsecuritynumbercombined} \\ 
4 & Tookavideoofyouenteringinyourpinatanatm & 92.67&0.475 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouenteringinyourPINatanATMcombined} \\ 
5 & Tookaphotoofyouunclothed & 92.59&0.4563 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookaphotoofyouunclothedcombined} \\ 
6 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 91.39&0.5567 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined} \\ 
7 & Learnedyourusernameandpasswordforwebsites & 89.55&0.6167 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourusernameandpasswordforwebsitescombined} \\ 
8 & Learnedyourcreditcardinformation & 88.98&0.5625 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourcreditcardinformationcombined} \\ 
9 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 88.41&0.5281 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined} \\ 
10 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 87.5&0.6041 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofyou(withaninward-facingcamera)athomecombined} \\ 
11 & Recordedyourworkconversations & 86.82&0.7602 & \includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourworkconversationscombined} \\ 
12 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 85.53&0.6227 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouenteringinadigitalpasscodetoalockeddoorcombined} \\ 
13 & Recordedyourphoneconversations & 85.16&0.6079 & \includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourphoneconversationscombined} \\ 
14 & Learnedhowmuchmoneyyouhave & 84.44&0.6142 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchmoneyyouhavecombined} \\ 
15 & Tookavideoofyouintoxicated & 83.21&0.721 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouintoxicatedcombined} \\ 
16 & Learnedwhenandhowmuchyouhavesex & 81.95&0.8185 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowmuchyouhavesexcombined} \\ 
17 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 81.05&0.5999 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofyou(withaninward-facingcamera)athomecombined} \\ 
18 & Tookaphotoofyouintoxicated & 78.95&0.8247 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookaphotoofyouintoxicatedcombined} \\ 
19 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 78.76&0.8525 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofyou(withaninward-facingcamera)atrandomcombined} \\ 
20 & Recordedyourpassingconversations & 78.13&0.8279 & \includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourpassingconversationscombined} \\ 
21 & Learnedwhatmedicalconditionsyouhave & 77.7&0.8609 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatmedicalconditionsyouhavecombined} \\ 
22 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 76.19&0.588 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofyou(withaninward-facingcamera)atrandomcombined} \\ 
23 & Tookavideoofyouoff-Guard & 76.0&0.6274 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouoff-guardcombined} \\ 
24 & Tookphotosatwork(Withanoutward-Facingcamera) & 74.62&0.9035 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosatwork(withanoutward-facingcamera)combined} \\ 
25 & Learnedyourusernameforwebsites & 73.44&0.8255 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourusernameforwebsitescombined} \\ 
26 & Learnedyouraddress & 72.61&0.8625 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyouraddresscombined} \\ 
27 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.55&0.7037 & \includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedaudiorecordingsyoumadeonyourdevicecombined} \\ 
28 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.55&0.7037 & \includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedaudiorecordingsyoumadeonyourdevicecombined} \\ 
29 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 71.81&0.9016 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosyoudownloadedfromtheinternetsavedonyourdevicecombined} \\ 
30 & Sharedphotosotherssenttoyousavedonyourdevice & 71.63&1.034 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosotherssenttoyousavedonyourdevicecombined} \\ 
31 & Sharedvideosotherssentyousavedonyourdevice & 70.59&0.8123 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosotherssentyousavedonyourdevicecombined} \\ 
32 & Tookvideosatwork(Withanoutward-Facingcamera) & 70.54&0.899 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosatwork(withanoutward-facingcamera)combined} \\ 
33 & Learnedyourfingerprintsomehow & 70.12&0.8582 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourfingerprintsomehowcombined} \\ 
34 & Learnedwhenyouwerelyingnervousorstressed & 69.74&0.9105 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouwerelyingnervousorstressedcombined} \\ 
35 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 69.59&0.9115 & \includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyoutalkingtoyourself(makingvoicenotes)combined} \\ 
36 & Learnedwhatmedicationyoutake & 69.49&1.0051 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatmedicationyoutakecombined} \\ 
37 & Sharedvideosyoutookwhicharealreadyonyourdevice & 68.89&0.8849 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosyoutookwhicharealreadyonyourdevicecombined} \\ 
38 & Tookapictureofyoursignature & 68.07&0.8389 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookapictureofyoursignaturecombined} \\ 
39 & Learnedwhatwebsitesyougoto & 66.44&1.0101 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatwebsitesyougotocombined} \\ 
40 & Sharedphotosyoutookwhicharealreadyonyourdevice & 66.21&1.0243 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosyoutookwhicharealreadyonyourdevicecombined} \\ 
41 & Learnedwhereyoulivesomehow & 65.0&0.9769 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyoulivesomehowcombined} \\ 
42 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 63.51&0.9892 & \includegraphics[width = 2cm, height = 0.5cm]{tables/trackedwhereyouareveryaccurately(whereyouareinaroom)combined} \\ 
43 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 61.94&1.063 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
44 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 61.49&1.0048 & \includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosyoudownloadedontheinternetsavedonyourdevicecombined} \\ 
45 & Learnedwhenandhowmuchyouspendtimealone & 61.27&0.9872 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowmuchyouspendtimealonecombined} \\ 
46 & Trackedwhereyouare(Likeagps) & 61.24&1.0812 & \includegraphics[width = 2cm, height = 0.5cm]{tables/trackedwhereyouare(likeaGPS)combined} \\ 
47 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 61.04&0.9495 & \includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
48 & Learnedwhereyouarecurrentlygoing & 60.87&0.9701 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyouarecurrentlygoingcombined} \\ 
49 & Recordedthesoundaroundyou & 60.45&0.9358 & \includegraphics[width = 2cm, height = 0.5cm]{tables/recordedthesoundaroundyoucombined} \\ 
50 & Learnedwhoyouwerespendingtimewith & 60.0&1.128 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhoyouwerespendingtimewithcombined} \\ 
51 & Learnedwhereyouworksomehow & 58.09&1.1581 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyouworksomehowcombined} \\ 
52 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 54.4&1.2884 & \includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)combined} \\ 
53 & Learnedhowmuchyouuseyourphone & 51.95&1.2246 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchyouuseyourphonecombined} \\ 
54 & Learnedwhatproductsyoubuy & 50.0&1.0937 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatproductsyoubuycombined} \\ 
55 & Learnedwhenyouaresickorhealthy & 48.17&1.271 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouaresickorhealthycombined} \\ 
56 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 46.98&1.1241 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowcloseyouaretootherpeopleyouinteractwithcombined} \\ 
57 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 46.81&1.3063 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturecombined} \\ 
58 & Learnedhowmuchyouuseyourcomputer & 44.93&1.1576 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchyouuseyourcomputercombined} \\ 
59 & Learnedwhenhowmuchandwhatyouareeating & 42.86&1.2707 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenhowmuchandwhatyouareeatingcombined} \\ 
60 & Learnedyourname & 42.54&1.3974 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyournamecombined} \\ 
61 & Learnedwhenandhowwellyouaresleeping & 40.56&1.3424 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowwellyouaresleepingcombined} \\ 
62 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 40.51&1.2671 & \includegraphics[width = 2cm, height = 0.5cm]{tables/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\ 
63 & Learnedwhenhowandhowmuchyouexercise & 38.66&1.2567 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenhowandhowmuchyouexercisecombined} \\ 
64 & Learnedwhenyouwerehappyorhavingfun & 34.75&1.2734 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouwerehappyorhavingfuncombined} \\ 
65 & Learnedwhattelevisionshowsyouwatch & 30.2&1.4009 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhattelevisionshowsyouwatchcombined} \\ 
66 & Learnedwhenyouarebusyorinterruptible & 29.5&1.2648 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouarebusyorinterruptiblecombined} \\ 
67 & Copiedanduploadedmusicfromyourdevice & 28.06&1.4257 & \includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedmusicfromyourdevicecombined} \\ 
68 & Learnedyourheartrate & 27.5&1.4026 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourheartratecombined} \\ 
69 & Learnedyourage & 24.29&1.4299 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyouragecombined} \\ 
70 & Learnedthelanguageyouwerespeaking & 15.86&1.485 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedthelanguageyouwerespeakingcombined} \\ 
71 & Learnedyourgender & 15.0&1.4577 & \includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourgendercombined} \\ 
\end{longtable}

\end{document}