### 2015-03-03 recipient.tex

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| r | p{1.5cm} | p{1cm} | p{1cm} | p{2cm} |}
\hline
Rank & Recipient & VUR & Variance & Distribution\\
\hline
1& WORKCONTACTS & 75.16 & 0.879886356298& \includegraphics[width = 2cm, height = 2cm]{tables/recipient_work}\\ 
2& PUBLIC & 72.41 & 0.96322375386& \includegraphics[width = 2cm, height = 2cm]{tables/recipient_pub}\\ 
3& FRIENDS & 69.47 & 1.03749698795 & \includegraphics[width = 2cm, height = 2cm]{tables/recipient_friends}\\ 
4& ALL & 64.78 & 1.1105854312& \includegraphics[width = 2cm, height = 2cm]{tables/recipient_total}\\ 
5& APPSERVER & 42.28 & 1.32348026053& \includegraphics[width = 2cm, height = 2cm]{tables/recipient_app}\\ 
6\hline
\end{tabular}
\end{center}
\end{table}