### 2015-03-03 benefit.tex
directory change in risk/benefit
title

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| p{2cm} | p{1cm} | p{1cm} | p{1cm} | c |}
\hline
Technology & Q1 &  Median & Q3 & Distribution  \\ 
\hline
Gender Detectionben & 10.0 & 10.0 & 15.0 & \includegraphics[width = 2cm, height = 0.5cm]{genderdetectionben} \\ 
Age Detectionben & 12.0 & 10.0 & 22.0 & \includegraphics[width = 2cm, height = 0.5cm]{agedetectionben} \\ 
Discreet Microphoneben & 15.0 & 10.0 & 20.0 & \includegraphics[width = 2cm, height = 0.5cm]{discreetmicrophoneben} \\ 
Cubetasticben & 15.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{cubetasticben} \\ 
Fitness Trackersben & 18.5 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{fitnesstrackersben} \\ 
Voice Based Emotion Detectionben & 20.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{voicebasedemotiondetectionben} \\ 
Facial Detectionben & 20.0 & 10.0 & 34.0 & \includegraphics[width = 2cm, height = 0.5cm]{facialdetectionben} \\ 
Discreet Video Cameraben & 20.0 & 15.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{discreetvideocameraben} \\ 
Google Glassben & 20.0 & 12.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{googleglassben} \\ 
Smartwatchesben & 20.0 & 10.0 & 35.0 & \includegraphics[width = 2cm, height = 0.5cm]{smartwatchesben} \\ 
Motorcycle & 20.0 & 12.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{MotorcycleBenefit} \\ 
Handgun & 20.0 & 10.0 & 30.0 & \includegraphics[width = 2cm, height = 0.5cm]{HandgunBenefit} \\ 
Facial Recognitionben & 22.0 & 12.5 & 42.5 & \includegraphics[width = 2cm, height = 0.5cm]{facialrecognitionben} \\ 
Lawnmower & 24.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{LawnmowerBenefit} \\ 
Speech To Textben & 25.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{speechtotextben} \\ 
Voice Recognitionben & 25.0 & 15.0 & 40.0 & \includegraphics[width = 2cm, height = 0.5cm]{voicerecognitionben} \\ 
Language Detectionben & 35.0 & 15.0 & 60.0 & \includegraphics[width = 2cm, height = 0.5cm]{languagedetectionben} \\ 
Heart Rate Detectionben & 40.0 & 26.0 & 65.0 & \includegraphics[width = 2cm, height = 0.5cm]{heartratedetectionben} \\ 
Location Trackingben & 40.0 & 20.0 & 70.0 & \includegraphics[width = 2cm, height = 0.5cm]{locationtrackingben} \\ 
Emailben & 50.0 & 29.0 & 77.5 & \includegraphics[width = 2cm, height = 0.5cm]{emailben} \\ 
Smartphonesben & 50.0 & 30.0 & 75.0 & \includegraphics[width = 2cm, height = 0.5cm]{smartphonesben} \\ 
Laptopsben & 60.0 & 40.0 & 80.0 & \includegraphics[width = 2cm, height = 0.5cm]{laptopsben} \\ 
Internetben & 65.0 & 45.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{internetben} \\ 
Electricity & 88.0 & 50.0 & 100.0 & \includegraphics[width = 2cm, height = 0.5cm]{ElectricityBenefit} \\ 
Ben & 100.0 & nan & nan & \includegraphics[width = 2cm, height = 0.5cm]{ben} \\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10}
\end{center}
\end{table}