### 2015-03-03 tex-inputs/long-tables/collapsedstd.tex

\documentclass[a4paper,12pt]{article}
\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\begin{document}
\begin{longtable}{| p{0.5cm} | p{7cm} | p{1cm} |p{1cm} | c |}

\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} &  \multicolumn{1}{c|}{\textbf{Standard Deviation}} &\multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \multicolumn{1}{|c|}{\textbf{Rank}} & \multicolumn{1}{c|}{\textbf{Question}} & \multicolumn{1}{c|}{\textbf{VUR}} & \multicolumn{1}{c|}{\textbf{Standard Deviation}} & \multicolumn{1}{c|}{\textbf{Distribution}} \\ \hline 
\endhead

\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
\endfoot
\hline 
\endlastfoot
1 & Learnedthelanguageyouwerespeaking & 15.86 & 1.485&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedthelanguageyouwerespeakingcombined} \\ 
2 & Learnedyourgender & 15.0 & 1.4577&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourgendercombined} \\ 
3 & Learnedyourage & 24.29 & 1.4299&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyouragecombined} \\ 
4 & Copiedanduploadedmusicfromyourdevice & 28.06 & 1.4257&\includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedmusicfromyourdevicecombined} \\ 
5 & Learnedyourheartrate & 27.5 & 1.4026&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourheartratecombined} \\ 
6 & Learnedwhattelevisionshowsyouwatch & 30.2 & 1.4009&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhattelevisionshowsyouwatchcombined} \\ 
7 & Learnedyourname & 42.54 & 1.3974&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyournamecombined} \\ 
8 & Learnedwhenandhowwellyouaresleeping & 40.56 & 1.3424&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowwellyouaresleepingcombined} \\ 
9 & Learnedhowyouwerefeelingbasedonheartratebreathingandortemperature & 46.81 & 1.3063&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturecombined} \\ 
10 & Copiedanduploadedsoundssavedonyourdevice(Notificationnoisesetc) & 54.4 & 1.2884&\includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)combined} \\ 
11 & Learnedwhenyouwerehappyorhavingfun & 34.75 & 1.2734&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouwerehappyorhavingfuncombined} \\ 
12 & Learnedwhenyouaresickorhealthy & 48.17 & 1.271&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouaresickorhealthycombined} \\ 
13 & Learnedwhenhowmuchandwhatyouareeating & 42.86 & 1.2707&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenhowmuchandwhatyouareeatingcombined} \\ 
14 & Scannedyoureyetolearnyoureyepatterns(Foreyetracking) & 40.51 & 1.2671&\includegraphics[width = 2cm, height = 0.5cm]{tables/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\ 
15 & Learnedwhenyouarebusyorinterruptible & 29.5 & 1.2648&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouarebusyorinterruptiblecombined} \\ 
16 & Learnedwhenhowandhowmuchyouexercise & 38.66 & 1.2567&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenhowandhowmuchyouexercisecombined} \\ 
17 & Learnedhowmuchyouuseyourphone & 51.95 & 1.2246&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchyouuseyourphonecombined} \\ 
18 & Learnedwhereyouworksomehow & 58.09 & 1.1581&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyouworksomehowcombined} \\ 
19 & Learnedhowmuchyouuseyourcomputer & 44.93 & 1.1576&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchyouuseyourcomputercombined} \\ 
20 & Learnedwhoyouwerespendingtimewith & 60.0 & 1.128&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhoyouwerespendingtimewithcombined} \\ 
21 & Learnedhowcloseyouaretootherpeopleyouinteractwith & 46.98 & 1.1241&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowcloseyouaretootherpeopleyouinteractwithcombined} \\ 
22 & Learnedwhatproductsyoubuy & 50.0 & 1.0937&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatproductsyoubuycombined} \\ 
23 & Trackedwhereyouare(Likeagps) & 61.24 & 1.0812&\includegraphics[width = 2cm, height = 0.5cm]{tables/trackedwhereyouare(likeaGPS)combined} \\ 
24 & Tookphotosofpeople(Withanoutward-Facingcamera)Atrandom & 61.94 & 1.063&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
25 & Sharedphotosotherssenttoyousavedonyourdevice & 71.63 & 1.034&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosotherssenttoyousavedonyourdevicecombined} \\ 
26 & Sharedphotosyoutookwhicharealreadyonyourdevice & 66.21 & 1.0243&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosyoutookwhicharealreadyonyourdevicecombined} \\ 
27 & Learnedwhatwebsitesyougoto & 66.44 & 1.0101&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatwebsitesyougotocombined} \\ 
28 & Learnedwhatmedicationyoutake & 69.49 & 1.0051&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatmedicationyoutakecombined} \\ 
29 & Sharedvideosyoudownloadedontheinternetsavedonyourdevice & 61.49 & 1.0048&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosyoudownloadedontheinternetsavedonyourdevicecombined} \\ 
30 & Trackedwhereyouareveryaccurately(Whereyouareinaroom) & 63.51 & 0.9892&\includegraphics[width = 2cm, height = 0.5cm]{tables/trackedwhereyouareveryaccurately(whereyouareinaroom)combined} \\ 
31 & Learnedwhenandhowmuchyouspendtimealone & 61.27 & 0.9872&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowmuchyouspendtimealonecombined} \\ 
32 & Learnedwhereyoulivesomehow & 65.0 & 0.9769&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyoulivesomehowcombined} \\ 
33 & Learnedwhereyouarecurrentlygoing & 60.87 & 0.9701&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhereyouarecurrentlygoingcombined} \\ 
34 & Tookvideosofpeople(Withanoutward-Facingcamera)Atrandom & 61.04 & 0.9495&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
35 & Recordedthesoundaroundyou & 60.45 & 0.9358&\includegraphics[width = 2cm, height = 0.5cm]{tables/recordedthesoundaroundyoucombined} \\ 
36 & Recordedyoutalkingtoyourself(Makingvoicenotes) & 69.59 & 0.9115&\includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyoutalkingtoyourself(makingvoicenotes)combined} \\ 
37 & Learnedwhenyouwerelyingnervousorstressed & 69.74 & 0.9105&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenyouwerelyingnervousorstressedcombined} \\ 
38 & Tookphotosatwork(Withanoutward-Facingcamera) & 74.62 & 0.9035&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosatwork(withanoutward-facingcamera)combined} \\ 
39 & Sharedphotosyoudownloadedfromtheinternetsavedonyourdevice & 71.81 & 0.9016&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedphotosyoudownloadedfromtheinternetsavedonyourdevicecombined} \\ 
40 & Tookvideosatwork(Withanoutward-Facingcamera) & 70.54 & 0.899&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosatwork(withanoutward-facingcamera)combined} \\ 
41 & Sharedvideosyoutookwhicharealreadyonyourdevice & 68.89 & 0.8849&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosyoutookwhicharealreadyonyourdevicecombined} \\ 
42 & Learnedyouraddress & 72.61 & 0.8625&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyouraddresscombined} \\ 
43 & Learnedwhatmedicalconditionsyouhave & 77.7 & 0.8609&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhatmedicalconditionsyouhavecombined} \\ 
44 & Learnedyourfingerprintsomehow & 70.12 & 0.8582&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourfingerprintsomehowcombined} \\ 
45 & Tookphotosofyou(Withaninward-Facingcamera)Atrandom & 78.76 & 0.8525&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofyou(withaninward-facingcamera)atrandomcombined} \\ 
46 & Tookapictureofyoursignature & 68.07 & 0.8389&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookapictureofyoursignaturecombined} \\ 
47 & Recordedyourpassingconversations & 78.13 & 0.8279&\includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourpassingconversationscombined} \\ 
48 & Learnedyourusernameforwebsites & 73.44 & 0.8255&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourusernameforwebsitescombined} \\ 
49 & Tookaphotoofyouintoxicated & 78.95 & 0.8247&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookaphotoofyouintoxicatedcombined} \\ 
50 & Learnedwhenandhowmuchyouhavesex & 81.95 & 0.8185&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedwhenandhowmuchyouhavesexcombined} \\ 
51 & Sharedvideosotherssentyousavedonyourdevice & 70.59 & 0.8123&\includegraphics[width = 2cm, height = 0.5cm]{tables/sharedvideosotherssentyousavedonyourdevicecombined} \\ 
52 & Tookaphotoofyouoff-Guard & 72.55 & 0.769&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookaphotoofyouoff-guardcombined} \\ 
53 & Recordedyourworkconversations & 86.82 & 0.7602&\includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourworkconversationscombined} \\ 
54 & Tookavideoofyouintoxicated & 83.21 & 0.721&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouintoxicatedcombined} \\ 
55 & Copiedanduploadedaudiorecordingsyoumadeonyourdevice & 72.55 & 0.7037&\includegraphics[width = 2cm, height = 0.5cm]{tables/copiedanduploadedaudiorecordingsyoumadeonyourdevicecombined} \\ 
56 & Tookavideoofyouoff-Guard & 76.0 & 0.6274&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouoff-guardcombined} \\ 
57 & Tookavideoofyouenteringinadigitalpasscodetoalockeddoor & 85.53 & 0.6227&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouenteringinadigitalpasscodetoalockeddoorcombined} \\ 
58 & Learnedyourusernameandpasswordforwebsites & 89.55 & 0.6167&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourusernameandpasswordforwebsitescombined} \\ 
59 & Learnedhowmuchmoneyyouhave & 84.44 & 0.6142&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedhowmuchmoneyyouhavecombined} \\ 
60 & Recordedyourphoneconversations & 85.16 & 0.6079&\includegraphics[width = 2cm, height = 0.5cm]{tables/recordedyourphoneconversationscombined} \\ 
61 & Tookphotosofyou(Withaninward-Facingcamera)Athome & 87.5 & 0.6041&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookphotosofyou(withaninward-facingcamera)athomecombined} \\ 
62 & Tookvideosofyou(Withaninward-Facingcamera)Athome & 81.05 & 0.5999&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofyou(withaninward-facingcamera)athomecombined} \\ 
63 & Tookvideosofyou(Withaninward-Facingcamera)Atrandom & 76.19 & 0.588&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookvideosofyou(withaninward-facingcamera)atrandomcombined} \\ 
64 & Learnedyourcreditcardinformation & 88.98 & 0.5625&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourcreditcardinformationcombined} \\ 
65 & Tookanincriminatingphotoofyoudoingsomethingembarrassing & 91.39 & 0.5567&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined} \\ 
66 & Tookanincriminatingvideoofyoudoingsomethingembarrassing & 88.41 & 0.5281&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined} \\ 
67 & Tookavideoofyouenteringinyourpinatanatm & 92.67 & 0.475&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouenteringinyourPINatanATMcombined} \\ 
68 & Tookaphotoofyouunclothed & 92.59 & 0.4563&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookaphotoofyouunclothedcombined} \\ 
69 & Learnedyourbankaccountinformation & 95.91 & 0.3539&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyourbankaccountinformationcombined} \\ 
70 & Tookavideoofyouunclothed & 95.97 & 0.3115&\includegraphics[width = 2cm, height = 0.5cm]{tables/tookavideoofyouunclothedcombined} \\ 
71 & Learnedyoursocialsecuritynumber & 94.84 & 0.26&\includegraphics[width = 2cm, height = 0.5cm]{tables/learnedyoursocialsecuritynumbercombined} \\ 
\end{longtable}

\end{document}