### 2015-03-04 paper.tex
got rid of longtables madness

\documentclass{acm_proc_article-sp}
\usepackage{color}
\usepackage[hyphens]{url}
%\usepackage[cm]{fullpage}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{enumitem}

\def\etal{{\it et al.~}}
\newenvironment{packed_enum}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}
\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\begin{document}
\input{tex-inputs/title.tex}
\input{tex-inputs/abstract.tex}
\input{tex-inputs/intro.tex}
\input{tex-inputs/related_work.tex} 
\input{tex-inputs/methodology.tex}

\input{tex-inputs/results-intro.tex}
\input{tex-inputs/results-type.tex}
\input{tex-inputs/results-recipient.tex}
\input{tex-inputs/results-typeandrecipient.tex}
\input{tex-inputs/results-devices.tex}
\input{tex-inputs/results-demographics.tex}
\input{tex-inputs/results-regression.tex}
\input{tex-inputs/results-techrank.tex}
\input{tex-inputs/results-openended.tex}

\input{tex-inputs/discussion.tex} 
\input{tex-inputs/conclusion.tex}

%\input{tables/collapsed.tex}
%\input{ranks.tex}
\input{tex-inputs/collapsed.tex}
\input{tex-inputs/VURcombined.tex}
\input{tex-inputs/risk.tex}
\input{tex-inputs/benefit.tex}
%\input{tex-inputs/riskbenefit.tex}
%\input{tex-inputs/appendix/riskbenefit-table.tex}
%\input{tex-inputs/appendix/VURcombined.tex}

\bibliographystyle{abbrv}
\bibliography{bibliography.bib} 
\input{tex-inputs/appendixes.tex}

\end{document}