### 2015-03-06 tex-inputs/results-intro.tex
Added clarifications

\section{Results}
In this section, we present participants' responses to the various data-sharing scenarios, and how data type, data recipient, and device contributed to their risk perceptions. Next, we discuss participants' risk/benefit assessment of various new technologies relative to well-established technologies. We conclude the section with participants' self-reported concerns about the biggest risks in owning wearable devices.

\subsection{Concern Factors}
Many factors impact participants' concern levels for each scenario: the data recipient, the data type, and whether or not the scenario occurred on a wearable or a smartphone. We analyze each factor individually, as well as present a statistical model of participants' concerns as a function of all of the factors, including demographic traits.