### 2015-03-07 tex-inputs/collapsed.tex

%!TEX root = ../paper.tex

%\onecolumn
\section{Concern Factors}
\label{sec:concerns-appendix} 
We show the full, fine-grained results of our survey in this appendix. This includes how participants had ranked each technology in response to the Fischhoff-style questions, the VUR rates for all seventy-two data types, across all recipients and by recipient, and the details of the full regression models used in our analyses.

\begin{table*}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | r | r | r |}
\hline
Rank & Question & VUR & $\sigma$ & Distribution \\
\hline
1 & video of you unclothed & 95.97 & 0.31 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouunclothedcombined} \\ 
2 & bank account information & 95.91 & 0.35 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourbankaccountinformationcombined} \\ 
3 & social security number & 94.84 & 0.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyoursocialsecuritynumbercombined} \\ 
4 & video entering in a PIN at an ATM & 92.67 & 0.48 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouenteringinyourPINatanATMcombined} \\ 
5 & photo of you unclothed & 92.59 & 0.45 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookaphotoofyouunclothedcombined} \\ 
6 & photo of you that is very embarrassing & 91.39 & 0.56 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined} \\ 
7 & username and password for websites & 89.55 & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourusernameandpasswordforwebsitescombined} \\ 
8 & credit card information & 88.98 & 0.56 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourcreditcardinformationcombined} \\ 
9 & video of you that is very embarrassing & 88.41 & 0.53 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined} \\ 
10 & photo of you at home & 87.5 &0.60 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookphotosofyou(withaninward-facingcamera)athomecombined} \\ 
11 & audio recording of work conversations & 86.82 & 0.76 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/recordedyourworkconversationscombined} \\ 
12 & video of entering in a passcode to a door & 85.53 & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouenteringinadigitalpasscodetoalockeddoorcombined} \\ 
13 & audio recording of phone conversations & 85.16 & 0.61 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/recordedyourphoneconversationscombined} \\ 
14 & amount of money you have & 84.44 & 0.61 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedhowmuchmoneyyouhavecombined} \\ 
15 & video of you intoxicated & 83.21 & 0.72 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouintoxicatedcombined} \\ 
16 & when you have sex & 81.95 & 0.82 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenandhowmuchyouhavesexcombined} \\ 
17 & how much debt you have & 81.12  & 0.54 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouintoxicatedcombined} \\ %learnedhowmuchdebtyouhavecombined
18 & video of you at home & 81.05 & 0.60 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookvideosofyou(withaninward-facingcamera)athomecombined} \\ 
19 & photo of you intoxicated & 78.95 & 0.82 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookaphotoofyouintoxicatedcombined} \\ 
20 & photo of you at random & 78.76 & 0.85 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookphotosofyou(withaninward-facingcamera)atrandomcombined} \\ 
21 & audio recording of conversations & 78.13 & 0.83 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/recordedyourpassingconversationscombined} \\ 
22 & medical conditions & 77.7 & 0.86 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhatmedicalconditionsyouhavecombined} \\ 
23 & video of you at random & 76.19 & 0.59 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookvideosofyou(withaninward-facingcamera)atrandomcombined} \\ 
24 & video of you off-guard & 76.0 & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookavideoofyouoff-guardcombined} \\ 
25 & photo of your work or workplace & 74.62 & 0.90 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookphotosatwork(withanoutward-facingcamera)combined} \\ 
26 & username for websites & 73.44 & 0.83 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourusernameforwebsitescombined} \\ 
27 & address & 72.61 & 0.86 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyouraddresscombined} \\ 
28 & audio recording you captured & 72.55  & 0.70 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/copiedanduploadedaudiorecordingsyoumadeonyourdevicecombined} \\ 
29 & photo of you off-guard & 72.55 & 0.77 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookaphotoofyouoff-guardcombined} \\ 
30 & photo downloaded from internet & 71.81 & 0.90 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedphotosyoudownloadedfromtheinternetsavedonyourdevicecombined} \\ 
31 & photo others sent you & 71.63 & 1.03 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedphotosotherssenttoyousavedonyourdevicecombined} \\ 
32 & video others sent you & 70.59 & 0.81 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedvideosotherssentyousavedonyourdevicecombined} \\ 
33 & video of your work or workplace & 70.54 & 0.90 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookvideosatwork(withanoutward-facingcamera)combined} \\ 
34 & fingerprint & 70.12 & 0.86 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourfingerprintsomehowcombined} \\ 
35 & when you were lying nervous or stressed & 69.74 & 0.91 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenyouwerelyingnervousorstressedcombined} \\ 
36 & audio recording of you (voice notes) & 69.59 & 0.91 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/recordedyoutalkingtoyourself(makingvoicenotes)combined} \\ 
& vdots & & & \\
\hline
\end{tabular}
\caption{VUR for all data types, across all recipients.}
\label{full-vur-table}
\end{center}
\end{table*}

\begin{table*}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | r | r | r |}
\hline
Rank & Question & VUR & $\sigma$ & Distribution \\
\hline
& \vdots & & & \\
37 & medication taken & 69.49 & 1.01 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhatmedicationyoutakecombined} \\ 
38 & videos already on device & 68.89 & 0.88 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedvideosyoutookwhicharealreadyonyourdevicecombined} \\ 
39 & photo of your signature & 68.07 & 0.84 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookapictureofyoursignaturecombined} \\ 
40 & web history & 66.44 & 1.01 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhatwebsitesyougotocombined} \\ 
41 & photos taken on device & 66.21 & 1.02 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedphotosyoutookwhicharealreadyonyourdevicecombined} \\ 
42 & home address & 65.0 & 0.97 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhereyoulivesomehowcombined} \\ 
43 & fine-grained location tracking (+/- cm) & 63.51 & 0.99 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/trackedwhereyouareveryaccurately(whereyouareinaroom)combined} \\ 
44 & photo of people at random & 61.94 & 1.06 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookphotosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
45 & video downloaded from the internet & 61.49 & 1.00 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/sharedvideosyoudownloadedontheinternetsavedonyourdevicecombined} \\ 
46 & when you are alone & 61.27 & 0.99 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenandhowmuchyouspendtimealonecombined} \\ 
47 & location tracking (+/- m) & 61.24 & 1.08 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/trackedwhereyouare(likeaGPS)combined} \\ 
48 & videos of people at random & 61.04 & 0.95 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/tookvideosofpeople(withanoutward-facingcamera)atrandomcombined} \\ 
49 & where you are currently going & 60.87 & 0.97 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhereyouarecurrentlygoingcombined} \\ 
50 & recording of sound around you & 60.45 & 0.94 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/recordedthesoundaroundyoucombined} \\ 
51 & people you spend time with & 60.0 & 1.13 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhoyouwerespendingtimewithcombined} \\ 
52 & workplace address & 58.09 & 1.16 &\includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhereyouworksomehowcombined} \\ 
53 & sounds on device (notifications, etc) & 54.4 & 1.29 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/copiedanduploadedsoundssavedonyourdevice(notificationnoisesetc)combined} \\ 
54 & phone usage & 51.95 &1.22 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedhowmuchyouuseyourphonecombined} \\ 
55 & purchased products & 50.0 & 1.09 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhatproductsyoubuycombined} \\ 
56 & when you are sick or healthy & 48.17  & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenyouaresickorhealthycombined} \\ 
57 & how close you are to interacting people & 46.98 & 1.12 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedhowcloseyouaretootherpeopleyouinteractwithcombined} \\ 
58 & feelings (based on biometrics) & 46.81 & 1.31 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedhowyouwerefeelingbasedonheartratebreathingandortemperaturecombined} \\ 
59 & computer usage & 44.93 & 1.16 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedhowmuchyouuseyourcomputercombined} \\ 
60 & eating patterns & 42.86 & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenhowmuchandwhatyouareeatingcombined} \\ 
61 & name & 42.54 & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyournamecombined} \\ 
62 & sleeping patterns & 40.56 & 1.34 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenandhowwellyouaresleepingcombined} \\ 
63 & eye patterns (for eye tracking) & 40.51 & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\ 
64 & exercise patterns & 38.66 & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenhowandhowmuchyouexercisecombined} \\ 
65 & when you are happy or having fun & 34.75 & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenyouwerehappyorhavingfuncombined} \\ 
66 & television shows watched & 30.2 & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhattelevisionshowsyouwatchcombined} \\ 
67 & when you are busy or interruptible & 29.5 & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedwhenyouarebusyorinterruptiblecombined} \\ 
68 & music on device & 28.06 & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/copiedanduploadedmusicfromyourdevicecombined} \\ 
69 & heart rate & 27.5 & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourheartratecombined} \\ 
70 & age & 24.29 & 1.43 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyouragecombined} \\ 
71 & language spoken & 15.86 & 1.49 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedthelanguageyouwerespeakingcombined} \\ 
72 & gender & 15.0 & 1.46 &  \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/table-images/learnedyourgendercombined} \\ 
\hline
\end{tabular}
\caption{VUR for all data types, across all recipients.}
\label{full-vur-table}
\end{center}
\end{table*}