### 2015-04-13 tex-inputs/conclusion.tex

%!TEX root = ../paper.tex

\section{Conclusion}

Wearables are still in their infancy. Perceptions of situations and capabilities will change rapidly with advancements and increased exposure. However, there is not much work done to determine which threats in the emerging threat landscape are pertinent to focus on. Our survey of 1,784 Internet users, is the first large-scale study to investigate user-centric security and privacy concerns for wearable devices We contribute a comprehensive ranking of possible risks associated with wearable devices, across various recipients. Inspection of possible data concerns agree with previous studies of smartphone studies to find that video capture and financial data to be most sensitive. This suggests that these risks should be highlighted by platform developers. Various systems which detect and take actions for sensitive objects in photos and videos will be critical as wearables and other devices become more ubiquitous. We calibrate our results with mobile devices and existing technologies; additionally, we verify with an open-ended question and find that privacy and security are at the top of user's overall concerns, over fashion, safety, and health concerns. We hope that this work has given a comprehensive overview of user concerns with respect to wearables, and inform designs of future privacy and security controls for wearable devices.